---
title: "Leaf photosynthetic traits and N from field observations"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rbeni)
library(patchwork)
```

## Data

The data used here is this leaf photosynthetic traits and N dataset:

### Data reference

Dong, Ning, Prentice, Iain Colin, Wright, Ian, Wang, Han, Atkin,Owen, Bloomfield, Keith, Domingues, Tomas, Gleason, Sean, Maire, Vincent, Onoda, Yusuke, Poorter, Hendrik, & Smith, Nicholas. (2022). dataset for paper "Leaf nitrogen from the perspective of optimal plant function" (Version v1) [Data set]. Zenodo. https://doi.org/10.5281/zenodo.6831903

### Paper reference

Dong, N., Prentice, I. C., Wright, I. J., Wang, H., Atkin, O. K., Bloomfield, K. J., Domingues, T. F., Gleason, S. M., Maire, V., Onoda, Y., Poorter, H., & Smith, N. G. (2022). Leaf nitrogen from the perspective of optimal plant function. Journal of Ecology, 00, 1– 18. https://doi.org/10.1111/1365-2745.13967

```{r}
df <- read_csv("~/data/leafn_ning_dong/data_leafn_ning_dong.csv")
```

## Distribution of data

```{r, warning=FALSE}
gg1 <- df %>% 
  ggplot(aes(vcmax25_obs)) +
  geom_density() +
  labs(title = "Nmass")

gg2 <- df %>% 
  ggplot(aes(Narea)) +
  geom_density() +
  labs(title = "Narea")

gg3 <- df %>% 
  ggplot(aes(Nmass)) +
  geom_density() +
  labs(title = "Nmass")

gg4 <- df %>% 
  ggplot(aes(LMA)) +
  geom_density() +
  labs(title = "LMA")

(gg1 + gg2) /
  (gg3 + gg4)
```

Log-transform all variables to make them closer to normally distributed.

```{r, warning=FALSE}
df <- df %>% 
  mutate(log_vcmax25_obs = log(vcmax25_obs),
         log_Narea = log(Narea),
         log_Nmass = log(Nmass),
         log_LMA = log(LMA),
         )
```

```{r, warning=FALSE}
gg1 <- df %>% 
  ggplot(aes(log_vcmax25_obs)) +
  geom_density() +
  labs(title = "Nmass")

gg2 <- df %>% 
  ggplot(aes(log_Narea)) +
  geom_density() +
  labs(title = "Narea")

gg3 <- df %>% 
  ggplot(aes(log_Nmass)) +
  geom_density() +
  labs(title = "Nmass")

gg4 <- df %>% 
  ggplot(aes(log_LMA)) +
  geom_density() +
  labs(title = "LMA")

(gg1 + gg2) /
  (gg3 + gg4)
```

## Analysis vs soil C:N

```{r, warning=FALSE}
gg1 <- df %>% 
  ggplot(aes(soil_cn, log_vcmax25_obs)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg2 <- df %>% 
  ggplot(aes(soil_cn, log_Narea)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg3 <- df %>% 
  ggplot(aes(soil_cn, log_Nmass)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg4 <- df %>% 
  ggplot(aes(soil_cn, log_LMA)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg1 + gg2 + gg3 + gg4
```