---
title: "Experiments data analysis with MESI"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    #theme: paper
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

```{r, warning=FALSE}
library(dplyr)       # for working with ease
library(tidyr)       # for working with ease
library(stringr)
library(ggplot2)
library(readr)
library(metafor)     # for meta analysis
library(MAd)         # for meta analysis
```


# Read data and libraries

Load and interpret MESI data files.
```{r message=FALSE}
# ## second line provides units
# col_names <- names(read_csv("~/mesi-db_test/data/mesi_main.csv", n_max = 0))
# units <- names(read_csv("~/mesi-db_test/data/mesi_main.csv", n_max = 0, skip = 1))
# df_units <- tibble(column = col_names,
#                    units = units
#                    )
# df <- read_csv("~/mesi-db_test/data/mesi_main.csv", col_names = col_names, skip = 2)
df <- read_csv("~/mesi-db_test/data/mesi_main.csv")
```


# CO2 experiments

Select experiments from which we will use data.
```{r warning=FALSE}
df0 <- df %>%
  
  ## CO2-only manipulation (no other factors manipulated, no crossed manipulation)
  filter(treatment == "c") %>% 
  
  ## only experiments conducted in the field
  filter(experiment_type == "field") %>% 
  
  ## use only free air CO2 enrichment experiments and open-top chambers
  filter(fumigation_type %in% c("face", "otc"))
```


## Data overview

**Issues**

- Some experiments don't have a name, referred to by coordinates. Can they be identified to get an experiments name? (-> STUDENT)


## Data subsetting

Let's use just data from CO2-only experiments that lasted at least three years.
```{r}
use_exp <- df0 %>% 
  filter(!is.na(sampling_year)) %>% 
  group_by(exp) %>% 
  summarise(nyears = max(sampling_year)) %>% 
  filter(nyears >= 3) %>% 
  pull(exp)

# df1 <- df0 %>%
#   filter(exp %in% use_exp)

## xxx try: do not subset to minimum experiment length
df1 <- df0
```

<!-- ### Available variables -->
<!-- ```{r} -->
<!-- df1 %>% -->
<!--   group_by(response) %>% -->
<!--   summarise(n = n()) %>% -->
<!--   arrange(response) %>% -->
<!--   knitr::kable() -->
<!-- ``` -->

<!-- ### Available experiments -->
<!-- ```{r} -->
<!-- df1 %>% -->
<!--   group_by(exp) %>%  -->
<!--   summarise(n_obs = sum(rep_c, na.rm = TRUE)) %>%  -->
<!--   arrange(exp) %>%  -->
<!--   knitr::kable() -->
<!-- ``` -->

### Select variables

**Issues:**

- No data available for Narea (-> KEVIN)

  
```{r}
## create own variable names (myvar), grouping available variables
use_response <- c("asat", "vcmax", "jmax", "gpp", "gs", "leaf_n_area", "leaf_n_mass", "leaf_cn", "anpp", "agb", "leaf_biomass", "lai", "lai_max", "leaf_area", "bgb", "fine_root_biomass", "root_production", "fine_root_production", "root_shoot_ratio", "soil_no3-n", "soil_nh4-n", "soil_nh4", "soil_no3", "soil_solution_nh4", "soil_solution_no3", "root_n_uptake", "root_nh4_uptake", "root_no3_uptake")

df2 <- df1 %>% 

  filter(response %in% use_response) %>% 
  mutate(myvar = response) %>% 

  ## variables re-grouped by myself
  mutate(myvar = ifelse(myvar %in%  c("agb", "agb_coarse"),   "agb", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("bgb", "fine_root_biomass", "coarse_root_c_stock", "bgb_coarse"), "bgb", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("lai", "lai_max"), "lai", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("bgb", "fine_root_biomass"), "bgb", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("root_production", "fine_root_production", "coarse_root_production"),
                        "root_production", 
                        myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("root_n_uptake", "root_nh4_uptake", "root_no3_uptake"),
                        "n_uptake", 
                        myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("soil_no3-n", "soil_nh4-n", "soil_nh4", "soil_no3", "soil_solution_nh4", "soil_solution_no3"), 
                        "n_inorg", 
                        myvar))

use_vars <- unique(df2$myvar)
```

<!-- ### Available experiments by variables -->
<!-- ```{r} -->
<!-- agg_by_var <- function(varnam, df){ -->
<!--   df %>%  -->
<!--     filter(myvar == varnam) %>%  -->
<!--     group_by(exp) %>%  -->
<!--     summarise(n = sum(rep_c)) -->
<!-- } -->

<!-- tmp2 <- purrr::map(as.list(use_vars), ~agg_by_var(., df2)) -->
<!-- names(tmp2) <- use_vars -->
<!-- tmp2$root_shoot_ratio -->
<!-- ``` -->

## Analysis

Test analysis and plot of ANPP data. Calculate the response ratio of ANPP (mean and variance) for each experiment. To get that, we first need to calcuate the means and standard deviation for the ambient and elevated levels, pooling multiple measurements (years, sampling dates), each given with mean $\mu_i$, number $N_i$ (replicates/plots), and standard deviation $\sigma_i$ or standard error. For the function `metafor::escalc()`, we need standard deviations ($SD$). Calculate them for those rows where only standard errors $SE$ are given as:
$$
SD = SE \sqrt{N}
$$

Calculate `"ROM"` - the log transformed ratio of means (Hedges et al., 1999; Lajeunesse, 2011) for each observation pair (ambient and elevated).
```{r}
df3 <- df2 %>%
  
  ## keep only essential variables and drop rows containing missing values for essential variables
  select(id, exp, myvar, treatment, sampling_year, x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>% 

  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = x_t, sd1i = sd_t, n1i = rep_t, 
    m2i = x_c, sd2i = sd_c, n2i = rep_c, 
    data = ., 
    append = TRUE, 
    var.names = c("logr", "logr_var") 
    ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  ## get standard error
  mutate( logr_se = sqrt(logr_var) / sqrt(rep_t) )

# ## data reduction here!
# tmp3 <- purrr::map(as.list(use_vars), ~agg_by_var(., df3))
# names(tmp3) <- use_vars
# tmp3$root_shoot_ratio
```

Aggregate all measurements (multiple years, sampling dates and plots) by experiment (and response variable - although here only one) for meta analysis.
```{r}
df4 <- df3 %>% 
  
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  
  # re-create ID (common ID per experiment and response variable)
  select(-id) %>%
  mutate( id = paste(exp, myvar, sep = "_XXX_")) %>% 
  
  MAd::agg( 
    id = id, 
    es = logr, 
    var = logr_var,
    cor = 1.0, 
    method = "BHHR", 
    data = . 
    ) %>% 

  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  # separate ID again for ease of data use
  mutate( id = str_split(id, "_XXX_") ) %>%
  mutate( exp = purrr::map_chr(id, 1),
          myvar = purrr::map_chr(id, 2) ) %>%
  
  ## rename again
  select(exp, myvar, logr = es, logr_var = var) %>%

  ## add number of observations (sum of plots and repeated samplings)
  left_join(
    df3 %>%
      group_by(exp, myvar, treatment) %>%
      summarise(n_c = sum(rep_c), n_t = sum(rep_t)),
    by = c("exp", "myvar")
  ) %>% 
  
  ## get standard error. Verify if number available observations are identical
  ## for ambient and elevated. Use N from control here (n_c).
  mutate( logr_se = sqrt(logr_var) / sqrt(n_c) )

# tmp <- purrr::map(as.list(use_vars), ~agg_by_var(., df4 %>% rename(rep_c = n_c)))
# names(tmp) <- use_vars
# tmp$root_shoot_ratio
```

Aggregate log-ratios across multiple experiments, taking into account their respective variance and using the experiment identity as a grouping factor for random intercepts.
```{r}
source("~/lt_cn_review/R/analyse_meta.R")
out  <- purrr::map(as.list(use_vars), ~analyse_meta(df4 %>% rename(var = myvar), nam_target = .))
names(out) <- use_vars
df_box <- purrr::map_dfr(out, "df_box")
saveRDS(df_box, file = "../data/df_box_mesi_co2.rds")
```

## Final data size

Number of data points (plot-level measurements) per variable:
```{r}
df4 %>% 
  group_by(myvar) %>% 
  summarise(n_plots = sum(n_c), n_exp = n()) %>% 
  rename_("Variable"="myvar", "N plots"="n_plots", "N experiments"="n_exp") %>% 
  knitr::kable()
```

Available experiments for root:shoot ratio
```{r}
ivar <- "root_shoot_ratio"
df4 %>%
  filter(myvar == ivar) %>%
  select(exp, n_c) %>%
  knitr::kable()
```

Available experiments for inorganic N availability
```{r}
ivar <- "n_inorg"
df4 %>%
  filter(myvar == ivar) %>%
  select(exp, n_c) %>%
  knitr::kable()
```

Available experiments for N uptake
```{r}
ivar <- "n_uptake"
df4 %>%
  filter(myvar == ivar) %>%
  select(exp, n_c) %>%
  knitr::kable()
```

## Visualisations

Plot dots and my box.
```{r}
df4 %>%
  ## give it a nice order (for plotting)
  mutate(myvar = factor(myvar, levels = rev(c("asat", "gpp", "vcmax", "jmax", "gs", "leaf_n_area", "leaf_n_mass", "leaf_cn", "lai", "leaf_area", "leaf_biomass", "anpp", "agb", "root_production", "bgb", "root_shoot_ratio", "n_uptake", "n_inorg")))) %>% 
  
  ggplot( aes(x = myvar, y = logr)) +
  geom_jitter( color = rgb(0,0,0,0.5), aes( size = 1/logr_se ), position = position_jitter(w = 0.2, h = 0) ) +
  geom_crossbar( data = df_box, 
                 aes(x = var, y = middle, ymin = ymin, ymax = ymax), 
                 fill = "tomato", color = "black", alpha = 0.6, width = 0.5 
                 ) +
  geom_hline( yintercept = 0.0, size = 0.5 ) +
  labs(x = "Variable", y = "Log Response Ratio", size = expression(paste("Error"^{-1}))) +
  coord_flip() +
  # ylim(-1, 1) +
  labs(title = "MESI data", subtitle = "Response to eCO2")
```

**Findings:**

- Strong positive response of leaf-level assimilation under light saturation (`asat`) and GPP
- Small (non-significant) reduction in Vcmax, no change in Jmax
- Clear decline in leaf N (I think it's Nmass, but to be confirmed -> KEVIN)
- Large range in plant-level (correct: plant-level? -> KEVIN) total leaf area
- Positive response in leaf biomass, even stronger than in LAI, indicating increase in (ecosystem-mean) LAI (-> KEVIN: to confirm: leaf_biomass and LAI are both ecosystem-level, while leaf_area is plant-level?)
- Strong positive response in root production - strongest response among all variables and stronger than response in ANPP (even considering negative outlier)
- Stronger response in belowground biomass (`bgb`) than in aboveground biomass (`agb`) (but outlier in `agb`), Consistent with the positive response in root:shoot ratio (but large variation between experiments).
- Positive response in N uptake, but large variation between experiments.
- Decline in mineral N availability (but not significant)
- Strong (in experiment extraordinarily strong) reduction in stomatal conductance (`gs`).

Standard forest plots

```{r}
for (ivar in use_vars){
  print(ivar)
  try(metafor::forest(out[[ivar]]$modl))
}
# purrr::map(as.list(use_vars), ~try(metafor::forest(out[[.]]$modl)))
```


**Issues:**

- Find out in which experiments, belowground biomass (`bgb`) was actually measured and not estimated based on aboveground biomass (or other aboveground observations) and scaling assumptions (-> KEVIN, STUDENT).
- Negative outliers for `agb` and `root_production` and positive outliers for `anpp` (-> KEVIN, remove them?)
- Euroface and Aspenface are provided as separate experiments. And several (versions of?) "Aspenface" are treated as separate experiments. Should they be? How do they differ? (-> KEVIN)
- Very little data on several variables (N uptake, root:shoot ratio, inorganic N availability, BNPP, ...). Wasn't there more in GCME (-> Beni)? Can we get more from published experiments in general (-> STUDENT)? Show data that was available in GCME for these variables:

```{r}
## file generated by cnreview/dataanalysis.Rmd
df_gcme <- readRDS("~/lt_cn_review/data/df_c_agg.rds")
write_csv(df_gcme, file = "~/lt_cn_review/data/df_co2_gcme_usedbybeni.csv")
```

N uptake
```{r}
## GCME
df_gcme %>% 
  filter(my_varnam == "my_nup") %>% 
  select(exp_nam, my_varnam, n_plots)

## MESI
df4 %>% 
  filter(myvar == "n_uptake") %>% 
  select(exp, myvar, n_c)
```

Root:shoot ratio
```{r}
## GCME
df_gcme %>% 
  filter(my_varnam == "my_rootshootratio") %>% 
  select(exp_nam, my_varnam, n_plots)

## MESI
df4 %>% 
  filter(myvar == "n_uptake") %>% 
  select(exp, myvar, n_c)
```

Inorganic N
```{r}
## GCME
df_gcme %>% 
  filter(my_varnam == "my_navl") %>% 
  select(exp_nam, my_varnam, n_plots)

## MESI
df4 %>% 
  filter(myvar == "n_inorg") %>% 
  select(exp, myvar, n_c)
```

Data in GCME from:
```{r}
df_navl <- read_rds("~/lt_cn_review/data/df_navl.rds")
df_navl %>% select(prev_name, Data_type, Fumigation_type, Source_Reference)
```

## Couplings

Plot effect sizes for two different variables against each other, given that data is available for both variables for a given experiment.

### ANPP-root production

```{r}
## make a wide table with ANPP and BNPP
df4_anpp_bnpp <- df4 %>% 
  filter(myvar %in% c("anpp", "root_production")) %>% 
  select(exp, myvar, logr) %>%
  tidyr::spread( myvar, logr )

## add standard error 
df4_anpp_bnpp <- df4 %>% 
  filter(myvar %in% c("anpp", "root_production")) %>% 
  select(exp, myvar, logr_se) %>%
  tidyr::spread( myvar, logr_se ) %>% 
  rename(se_anpp = anpp, se_root_production = root_production) %>%
  right_join(df4_anpp_bnpp, by = "exp") %>% 
  mutate(se = se_anpp * se_root_production)

df4_anpp_bnpp %>% 
  ggplot(aes(x = anpp, y = root_production, label = exp)) +
  geom_point(aes(size =  1/se), color = "tomato") +
  xlim(0, 0.8) + ylim(0, 0.8) +
  geom_abline(linetype = "dotted") + 
  ggrepel::geom_text_repel(size = 3, point.padding = 0.5, segment.alpha = 0, color = "grey50") +
  labs(x = "Log RR of ANPP", y = "Log RR of root production", size = expression(paste("Error"^{-1}))) +
  theme_classic()
```

### AGB - BGB

```{r}
## make a wide table with ANPP and BNPP
df4_agb_bnpp <- df4 %>% 
  filter(myvar %in% c("agb", "bgb")) %>% 
  select(exp, myvar, logr) %>%
  tidyr::spread( myvar, logr )

## add standard error 
df4_agb_bnpp <- df4 %>% 
  filter(myvar %in% c("agb", "bgb")) %>% 
  select(exp, myvar, logr_se) %>%
  tidyr::spread( myvar, logr_se ) %>% 
  rename(se_agb = agb, se_bgb = bgb) %>%
  right_join(df4_agb_bnpp, by = "exp") %>% 
  mutate(se = se_agb * se_bgb)

df4_agb_bnpp %>% 
  ggplot(aes(x = agb, y = bgb, label = exp)) +
  geom_point(aes(size =  1/se), color = "tomato") +
  xlim(0, 0.6) + ylim(0, 0.6) +
  geom_abline(linetype = "dotted") + 
  ggrepel::geom_text_repel(size = 3, point.padding = 0.5, segment.alpha = 0, color = "grey50") +
  labs(x = "Log RR of AGB", y = "Log RR of BGB", size=expression(paste("Error"^{-1}))) +
  theme_classic()
```

**Issues:**

- SwissFACE Lolium: I seem to remember that there was a stronger stimulation of belowground than aboveground biomass. Why does this here suggest a much higher response in AGB than in BGB? In the SwissFACE experiment, ABG was determined from multiple cuts per year. Within each year, harvested (cut) biomass should be summed up, not averaged! Please check and correct: where sums are relevant, take sum over multiple samplings per year. (-> KEVIN, STUDENT)




# N fertilisation experiments

Select experiments from which we will use data.
```{r warning=FALSE}
df0 <- df %>%
  
  ## fertilisation experiments only
  filter(treatment == "f") %>% 
  
  ## only experiments conducted in the field
  filter(experiment_type == "field")
```

## Data subsetting

Same as above
```{r}
use_exp <- df0 %>% 
  filter(!is.na(sampling_year)) %>% 
  group_by(exp) %>% 
  summarise(nyears = max(sampling_year)) %>% 
  filter(nyears >= 3) %>% 
  pull(exp)

# df1 <- df0 %>%
#   filter(exp %in% use_exp)

## xxx try: do not subset to minimum experiment length
df1 <- df0
```

### Select variables
  
```{r}
## select variables only variables as used for CO2 analysis and not provided by Liang et al.
## create own variable names (myvar), grouping available variables
use_response <- c("anpp", "agb", "bgb", "fine_root_biomass", "root_production", "fine_root_production", "root_shoot_ratio", "soil_no3-n", "soil_nh4-n", "soil_nh4", "soil_no3", "soil_solution_nh4", "soil_solution_no3", "root_n_uptake", "root_nh4_uptake", "root_no3_uptake")

df2 <- df1 %>% 

  filter(response %in% use_response) %>% 
  mutate(myvar = response) %>% 

  ## variables re-grouped by myself
  ## variables re-grouped by myself
  mutate(myvar = ifelse(myvar %in%  c("agb", "agb_coarse"),   "agb", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("bgb", "fine_root_biomass", "coarse_root_c_stock", "bgb_coarse"), "bgb", myvar)) %>% 
  # mutate(myvar = ifelse(myvar %in%  c("lai", "lai_max"), "lai", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("bgb", "fine_root_biomass"), "bgb", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("root_production", "fine_root_production", "coarse_root_production"),
                        "root_production", 
                        myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("root_n_uptake", "root_nh4_uptake", "root_no3_uptake"),
                        "n_uptake", 
                        myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("soil_no3-n", "soil_nh4-n", "soil_nh4", "soil_no3", "soil_solution_nh4", "soil_solution_no3"), 
                        "n_inorg", 
                        myvar))
  
use_vars <- unique(df2$myvar)
```


## Analysis

Calculate `"ROM"` - the log transformed ratio of means (Hedges et al., 1999; Lajeunesse, 2011) for each observation pair (ambient and elevated).
```{r}
df3 <- df2 %>%
  
  ## keep only essential variables and drop rows containing missing values for essential variables
  select(id, exp, myvar, treatment,sampling_year, x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>% 

  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = x_t, sd1i = sd_t, n1i = rep_t, 
    m2i = x_c, sd2i = sd_c, n2i = rep_c, 
    data = ., 
    append = TRUE, 
    var.names = c("logr", "logr_var") 
    ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  ## get standard error
  mutate( logr_se = sqrt(logr_var) / sqrt(rep_t) )
```

Aggregate all measurements (multiple years, sampling dates and plots) by experiment (and response variable - although here only one) for meta analysis.
```{r}
df4 <- df3 %>% 
  
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  
  # re-create ID (common ID per experiment and response variable)
  select(-id) %>%
  mutate( id = paste(exp, myvar, sep = "_XXX_")) %>% 
  
  MAd::agg( 
    id = id, 
    es = logr, 
    var = logr_var,
    cor = 1.0, 
    method = "BHHR", 
    data = . 
    ) %>% 

  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  # separate ID again for ease of data use
  mutate( id = str_split(id, "_XXX_") ) %>%
  mutate( exp = purrr::map_chr(id, 1),
          myvar = purrr::map_chr(id, 2) ) %>%
  
  ## rename again
  select(exp, myvar, logr = es, logr_var = var) %>%

  ## add number of observations (sum of plots and repeated samplings)
  left_join(
    df3 %>%
      group_by(exp, myvar, treatment) %>%
      summarise(n_c = sum(rep_c), n_t = sum(rep_t)),
    by = c("exp", "myvar")
  ) %>% 
  
  ## get standard error. Verify if number available observations are identical
  ## for ambient and elevated. Use N from control here (n_c).
  mutate( logr_se = sqrt(logr_var) / sqrt(n_c) )
```

Aggregate log-ratios across multiple experiments, taking into account their respective variance and using the experiment identity as a grouping factor for random intercepts.
```{r}
source("~/lt_cn_review/R/analyse_meta.R")
out  <- purrr::map(as.list(use_vars), ~analyse_meta(df4 %>% rename(var = myvar), nam_target = .))
names(out) <- use_vars
df_box <- purrr::map_dfr(out, "df_box")
saveRDS(df_box, file = "../data/df_box_mesi_nfert.rds")
```

## Final data size

Number of data points (plot-level measurements) per variable:
```{r}
df4 %>% 
  group_by(myvar) %>% 
  summarise(n_plots = sum(n_c), n_exp = n()) %>% 
  rename_("Variable"="myvar", "N observations"="n_plots", "N experiments"="n_exp") %>% 
  knitr::kable()
```

Number of data points (plot-level measurements) per experiment:
```{r}
df4 %>% 
  group_by(exp) %>% 
  summarise(n_plots = sum(n_c), n_exp = n()) %>% 
  rename_("Experiment"="exp", "N observations"="n_plots", "N experiments"="n_exp") %>% 
  knitr::kable()
```

## Visualisations

Plot dots and my box.
```{r}
df4 %>%
  drop_na(myvar) %>% 
  
  ## give it a nice order (for plotting)
  mutate(myvar = factor(myvar, levels = rev(c("anpp", "agb", "root_production", "bgb", "root_shoot_ratio", "n_uptake", "n_inorg")))) %>% 

  ggplot( aes(x = myvar, y = logr)) +
  geom_jitter( color = rgb(0,0,0,0.5), 
               aes( size = 1/logr_se ), 
               position = position_jitter(w = 0.2, h = 0) 
               ) +
  geom_crossbar( data = df_box %>% drop_na(var), 
                 aes(x = var, y = middle, ymin = ymin, ymax = ymax),
                 fill = "tomato", color = "black", alpha = 0.6, width = 0.5 ) +
  geom_hline( yintercept = 0.0, size = 0.5 ) +
  labs(x = "Variable", y = "Log Response Ratio", size = expression(paste("Error"^{-1}))) +
  coord_flip() +
  labs(title = "MESI data", subtitle = "Response to N-fertilisation")
```

**Findings:**

- Stronger effect on aboveground biomass than on belowground biomass - this comes out very clearly thanks to lots of data. One outlier of a very negative `agb` response (-> STUDENT?)
- Reduction in root:shoot ratio - opposite to CO2 response and consistent with expectations from a functional balance.
- Unsurprisingly, inorganic soil N increases in some cases very strongly, but large variation between experiments.
- No significant change in root production. Is this reliable? Or just to difficult to measure?  

**Issues:**

- Get more data for root:shoot ratio from N-fertilisation experiments (-> Evan).
- Get more data for N-inorg. data from N-fertilisation experiments (-> Evan).

```{r}
for (ivar in use_vars){
  print(ivar)
  try(metafor::forest(out[[ivar]]$modl))
}
# purrr::map(as.list(use_vars), ~try(metafor::forest(out[[.]]$modl)))
```


## Couplings

Plot effect sizes for two different variables against each other, given that data is available for both variables for a given experiment.

### ANPP-root production

```{r}
## make a wide table with ANPP and BNPP
df4_anpp_bnpp <- df4 %>% 
  filter(myvar %in% c("anpp", "root_production")) %>% 
  select(exp, myvar, logr) %>%
  tidyr::spread( myvar, logr )

## add standard error 
df4_anpp_bnpp <- df4 %>% 
  filter(myvar %in% c("anpp", "root_production")) %>% 
  select(exp, myvar, logr_se) %>%
  tidyr::spread( myvar, logr_se ) %>% 
  rename(se_anpp = anpp, se_root_production = root_production) %>%
  right_join(df4_anpp_bnpp, by = "exp") %>% 
  mutate(se = se_anpp * se_root_production)

df4_anpp_bnpp %>% 
  ggplot(aes(x = anpp, y = root_production, label = exp)) +
  geom_point(aes(size =  1/se), color = "tomato") +
  xlim(0, 0.8) + ylim(0, 0.8) +
  geom_abline(linetype = "dotted") + 
  ggrepel::geom_text_repel(size = 3, point.padding = 0.5, segment.alpha = 0, color = "grey50") +
  labs(x = "Log RR of ANPP", y = "Log RR of root production", size = expression(paste("Error"^{-1}))) +
  theme_classic()
```

**Issues:**

- Not enough data. Drop this analysis.


### AGB - BGB

```{r}
## make a wide table with ANPP and BNPP
df4_agb_bnpp <- df4 %>% 
  filter(myvar %in% c("agb", "bgb")) %>% 
  select(exp, myvar, logr) %>%
  tidyr::spread( myvar, logr )

## add standard error 
df4_agb_bnpp <- df4 %>% 
  filter(myvar %in% c("agb", "bgb")) %>% 
  select(exp, myvar, logr_se) %>%
  tidyr::spread( myvar, logr_se ) %>% 
  rename(se_agb = agb, se_bgb = bgb) %>%
  right_join(df4_agb_bnpp, by = "exp") %>% 
  mutate(se = se_agb * se_bgb)

df4_agb_bnpp %>% 
  ggplot(aes(x = agb, y = bgb, label = exp)) +
  geom_point(aes(size =  1/se), color = "tomato") +
  xlim(0, 0.6) + ylim(0, 0.6) +
  geom_abline(linetype = "dotted") + 
  ggrepel::geom_text_repel(size = 3, point.padding = 0.5, segment.alpha = 0, color = "grey50") +
  labs(x = "Log RR of AGB", y = "Log RR of BGB", size=expression(paste("Error"^{-1}))) +
  theme_classic()
```

**Issues:**

- Very few data on belowground biomass response. No estimates available from other experiments, e.g., from NutNet (-> Evan)

<!-- # N fertilisation experiments -->

<!-- Select data based on our purpose. -->
<!-- ```{r} -->
<!-- df_sub <- df %>%  -->
<!--   # filter(str_detect(treatment, "f")) %>%  -->
<!--   filter(treatment == "f") %>%  -->
<!--   filter(response %in% c("vcmax", "leaf_n", "leaf_cn")) %>%  -->
<!--   filter(experiment_type %in% c("field", "greenhouse", "growth_chamber", "outdoor_chamber"))  # not "pot" -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_sub %>%  -->
<!--   select(id, exp, response, experiment_type, treatment, x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>%  -->
<!--   drop_na() %>%  -->
<!--   group_by(response) %>%  -->
<!--   summarise(n_data = n()) %>%  -->
<!--   left_join( -->
<!--     df_sub %>%  -->
<!--       select(exp, response) %>%  -->
<!--       drop_na() %>% -->
<!--       distinct() %>%  -->
<!--       group_by(response) %>%  -->
<!--       summarise(n_experiments = n()), -->
<!--     by = "response" -->
<!--   ) -->
<!-- ``` -->


<!-- ## Analysis -->



<!-- ## Comments -->

<!-- - Needs extensive explanation of each column -->
<!-- - No data for Narea or should this be derived from leaf_n and LMA? -->
<!-- - Standardised units per variable (response). Currently, values are given in different units. This could be post-processed quite easily. -->
<!-- - Units of `leaf_cn` is given as `"molar"` or `NA`. Please provide information about whether `NA` means gC/gN. -->
<!-- ```{r} -->
<!-- df %>%  -->
<!--   filter(response %in% c("leaf_n")) %>% -->
<!--   pull(x_units) %>%  -->
<!--   unique() -->

<!-- df %>%  -->
<!--   filter(response %in% c("vcmax")) %>% -->
<!--   pull(x_units) %>%  -->
<!--   unique() -->

<!-- df %>%  -->
<!--   filter(response %in% c("leaf_cn")) %>% -->
<!--   pull(x_units) -->
<!-- ``` -->
