---
title: "Experiments data analysis with Liang et al. 2020 data"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    #theme: paper
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

```{r, warning=FALSE}
library(tidyverse)   # for working with ease
library(metafor)     # for meta analysis
```


# Read data and libraries

The intention is to use the MESI db to perform a meta analysis of variables under N fertilisation:

- ...

Load data files
```{r}
# files <- list.files(path = "~/data/nfert_liang/data_csv/", full.names = TRUE)
# df <- purrr::map_dfr(as.list(files), ~read_csv(.))
# use_names <- c(
#   "variable",           "reference",          "lat",                "lon",                "mat",               
#   "map",                "ph_soil",            "species",            "family",             "genus",             
#   "photopathway",       "legume",             "managed",            "growthform",         "functionaltype",    
#   "broadleaf_conifer",  "evergreen_deciduous","ninorg_form",        "ndep_kg_ha_yr",      "duration_yr",       
#   "load",               "value_t",            "value_c",            "n_t",                "n_c",               
#   "sd_t",               "sd_c",               "ln_rr",              "v",                  "w",                 
#   "cluster",            "w_dash"
#   )
# names(df) <- use_names
# write_csv(df, file = "~/data/nfert_liang/nfert_liang_combined.csv")
  
# modify nfert_liang_combined.csv by hand to have nice column names.
df <- read_csv("~/data/nfert_liang/nfert_liang_combined.csv")

# use_vars <- c("Vcmax", "Narea", "Nmass", "LMA", "SLA")
use_vars <- unique(df$variable)
```

Make data (more) conform with MESI.
```{r}
df <- df %>% 
  mutate(db = "liang",
         treatment = "f") %>% 
  rename(citation = reference,
         response = variable,
         x_c = value_c,
         x_t = value_t,
         rep_c = n_c,
         rep_t = n_t,
         growth_form = growthform
         ) %>% 
  mutate(growth_form = ifelse("Woody", # Liang name
                              "woody", # MESI name
                              ifelse("Herb",      # Liang name
                                     "herbaceous" # MESI name
                                     )))

# ## growth forms in Liang and in MESI
# df$growthform %>% unique()
# df$growth_form %>% unique()
```

Missing info in Liang et al.:
- elevation (`elv`)


## Data overview

**Issues**

- Is information on fumigation type available in some way? I'd like to select data from FACE and open-top chambers.
- Some experiments don't have a name, referred to by coordinates.

## Data subsetting

Let's use just data for above specified variables.
```{r}
df2 <- df %>% 
  filter(response %in% use_vars)
```

## Analysis

Test analysis and plot of ANPP data. Calculate the response ratio of ANPP (mean and variance) for each experiment. To get that, we first need to calcuate the means and standard deviation for the ambient and elevated levels, pooling multiple measurements (years, sampling dates), each given with mean $\mu_i$, number $N_i$ (replicates/plots), and standard deviation $\sigma_i$ or standard error. For the function `metafor::escalc()`, we need standard deviations ($SD$). Calculate them for those rows where only standard errors $SE$ are given as:
$$
SD = SE \sqrt{N}
$$

Calculate `"ROM"` - the log transformed ratio of means (Hedges et al., 1999; Lajeunesse, 2011) for each observation pair (ambient and elevated).
```{r}
df3 <- df2 %>%
  mutate(id = 1:n()) %>% 
  filter(sd_c > 0, sd_t > 0) %>% 
  
  ## keep only essential variables and drop rows containing missing values for essential variables
  select(id, exp = citation, response, x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>% 
  drop_na() %>% 
  
  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = x_t, sd1i = sd_t, n1i = rep_t, 
    m2i = x_c, sd2i = sd_c, n2i = rep_c, 
    data = ., 
    append = TRUE, 
    var.names = c("logr", "logr_var") 
    ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  ## get standard error
  mutate( logr_se = sqrt(logr_var) / sqrt(rep_t) )
```

Aggregate all measurements (species, multiple years, sampling dates and plots) by experiment (and response variable - although here only one) for meta analysis.
```{r}
df4 <- df3 %>% 
  
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  
  # re-create ID (common ID per experiment and response variable)
  select(-id) %>%
  mutate( id = paste(exp, response, sep = "_XXX_")) %>% 
  
  MAd::agg( 
    id = id, 
    es = logr, 
    var = logr_var,
    cor = 1.0, 
    method = "BHHR", 
    data = . 
    ) %>% 

  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  # separate ID again for ease of data use
  mutate( id = str_split(id, "_XXX_") ) %>%
  mutate( exp = purrr::map_chr(id, 1),
          response = purrr::map_chr(id, 2) ) %>%
  
  ## rename again
  select(exp, response, logr = es, logr_var = var) %>%

  ## add number of observations (sum of plots and repeated samplings)
  left_join(
    df3 %>%
      group_by(exp, response) %>%
      summarise(n_c = sum(rep_c), n_t = sum(rep_t)),
    by = c("exp", "response")
  ) %>% 
  
  ## get standard error. Verify if number available observations are identical
  ## for ambient and elevated. Use N from control here (n_c).
  mutate( logr_se = sqrt(logr_var) / sqrt(n_c) )
```

Aggregate log-ratios across multiple experiments, taking into account their respective variance and using the experiment identity as a grouping factor for random intercepts.
```{r}
source("~/lt_cn_review/R/analyse_meta.R")
out  <- purrr::map(as.list(use_vars), ~analyse_meta(df4, nam_target = .))
names(out) <- use_vars
df_box <- purrr::map_dfr(out, "df_box")
```


## Visualisations

Plot dots and my box.
```{r}
df4 %>%
  ggplot( aes(x = response, y = logr)) +
  geom_jitter( color = rgb(0,0,0,0.3), aes( size = 1/logr_se ), position = position_jitter(w = 0.2, h = 0) ) +
  geom_crossbar( data = df_box, aes(x = response, y = middle, ymin = ymin, ymax = ymax), fill = "tomato", color = "tomato4", alpha = 0.6, width = 0.5 ) +
  geom_hline( yintercept = 0.0, size = 0.5, linetype = "dotted") +
  labs(x = "Variable", y = "Log Response Ratio", size = expression(paste("Error"^{-1}))) +
  coord_flip() +
  ylim(-1, 1)  +
  labs(title = "Liang et al. 2020 data")
```


```{r}
# Vcmax
metafor::forest(out$Vcmax$modl)

# Narea
metafor::forest(out$Narea$modl)

# Nmass
metafor::forest(out$Nmass$modl)

# LMA
metafor::forest(out$LMA$modl)

# SLA
metafor::forest(out$SLA$modl)
```


<!-- # N fertilisation experiments -->

<!-- Select data based on our purpose. -->
<!-- ```{r} -->
<!-- df_sub <- df %>%  -->
<!--   # filter(str_detect(treatment, "f")) %>%  -->
<!--   filter(treatment == "f") %>%  -->
<!--   filter(response %in% c("vcmax", "leaf_n", "leaf_cn")) %>%  -->
<!--   filter(experiment_type %in% c("field", "greenhouse", "growth_chamber", "outdoor_chamber"))  # not "pot" -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_sub %>%  -->
<!--   select(id, exp, response, experiment_type, treatment, x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>%  -->
<!--   drop_na() %>%  -->
<!--   group_by(response) %>%  -->
<!--   summarise(n_data = n()) %>%  -->
<!--   left_join( -->
<!--     df_sub %>%  -->
<!--       select(exp, response) %>%  -->
<!--       drop_na() %>% -->
<!--       distinct() %>%  -->
<!--       group_by(response) %>%  -->
<!--       summarise(n_experiments = n()), -->
<!--     by = "response" -->
<!--   ) -->
<!-- ``` -->


<!-- ## Analysis -->



<!-- ## Comments -->

<!-- - Needs extensive explanation of each column -->
<!-- - No data for Narea or should this be derived from leaf_n and LMA? -->
<!-- - Standardised units per variable (response). Currently, values are given in different units. This could be post-processed quite easily. -->
<!-- - Units of `leaf_cn` is given as `"molar"` or `NA`. Please provide information about whether `NA` means gC/gN. -->
<!-- ```{r} -->
<!-- df %>%  -->
<!--   filter(response %in% c("leaf_n")) %>% -->
<!--   pull(x_units) %>%  -->
<!--   unique() -->

<!-- df %>%  -->
<!--   filter(response %in% c("vcmax")) %>% -->
<!--   pull(x_units) %>%  -->
<!--   unique() -->

<!-- df %>%  -->
<!--   filter(response %in% c("leaf_cn")) %>% -->
<!--   pull(x_units) -->
<!-- ``` -->
