---
title: "Experiments data analysis with Liang et al. 2020 data"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    #theme: paper
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

```{r, warning=FALSE}
library(tidyverse)   # for working with ease
library(metafor)     # for meta analysis
library(MAd)         # for meta analysis
source("~/lt_cn_review/R/analyse_meta.R")
```


The Liang et al. data can be obtained as an Excel file from here: https://doi.org/10.1111/gcb.15071.
I first exported each excel tab to a separate CSV file and then combined them into a single big table. Respective code is commented out below and the combined file is read directly.
```{r}
# files <- list.files(path = "~/data/nfert_liang/data_csv/", full.names = TRUE)
# df <- purrr::map_dfr(as.list(files), ~read_csv(.))
# use_names <- c(
#   "variable",           "reference",          "lat",                "lon",                "mat",
#   "map",                "ph_soil",            "species",            "family",             "genus",
#   "photopathway",       "legume",             "managed",            "growthform",         "functionaltype",
#   "broadleaf_conifer",  "evergreen_deciduous","ninorg_form",        "ndep_kg_ha_yr",      "duration_yr",
#   "load",               "value_t",            "value_c",            "n_t",                "n_c",
#   "sd_t",               "sd_c",               "ln_rr",              "v",                  "w",
#   "cluster",            "w_dash"
#   )
# names(df) <- use_names
# write_csv(df, file = "~/data/nfert_liang/nfert_liang_combined.csv")
  
# modify nfert_liang_combined.csv by hand to have nice column names.
df <- read_csv("~/data/nfert_liang/nfert_liang_combined.csv")
```

Make data (more) conform with MESI.
```{r}
# ALL:
# use_vars <- c("Area", "Vcmax", "gs", "inst-WUE", "Narea", "Nmass", "SLA", "LMA", "LAI", "Leaf area per plant", "Leaf biomass")

use_vars <- c("Aarea", 
              "Vcmax", 
              "Narea", 
              "Nmass",
              "LAI",
              # "Leaf area per plant", 
              "Leaf biomass"
              )

df <- df %>% 
  mutate(db = "liang",
         treatment = "f") %>% 
  rename(citation = reference,
         response = variable,
         x_c = value_c,
         x_t = value_t,
         rep_c = n_c,
         rep_t = n_t,
         growth_form = growthform
         ) %>% 
  mutate(growth_form = ifelse("Woody", # Liang name
                              "woody", # MESI name
                              ifelse("Herb",      # Liang name
                                     "herbaceous" # MESI name
                                     ))) |> 
  
  ## to make it comparable, chose only data below 300 kg ha-1 yr-1
  filter(ndep_kg_ha_yr <= 300) |>
  
  ## Use only variables that were also used for the MESI-CO2 meta-analysis
  filter(response %in% use_vars)
```

Missing info in Liang et al.:
- elevation (`elv`)

Classify into N-fertilisation levels (l1, l2, l3).
```{r}
df <- df |> 
  mutate(nfert_lev = ifelse(ndep_kg_ha_yr < 100, 
                            "f1",
                            ifelse(ndep_kg_ha_yr < 200,
                                   "f2",
                                   "f3")))
```

Distribution of N treatment level
```{r eval=FALSE}
ggplot(data = df,
       aes(x = ndep_kg_ha_yr, y = ..density.., fill = nfert_lev)) +
  geom_histogram()
```

## Data overview

## Data subsetting

Let's use just data for above specified variables.
```{r}
df2 <- df %>% 
  filter(response %in% use_vars)
```

## Analysis

Test analysis and plot of ANPP data. Calculate the response ratio of ANPP (mean and variance) for each experiment. To get that, we first need to calcuate the means and standard deviation for the ambient and elevated levels, pooling multiple measurements (years, sampling dates), each given with mean $\mu_i$, number $N_i$ (replicates/plots), and standard deviation $\sigma_i$ or standard error. For the function `metafor::escalc()`, we need standard deviations ($SD$). Calculate them for those rows where only standard errors $SE$ are given as:
$$
SD = SE \sqrt{N}
$$

## Response ratios

Calculate `"ROM"` - the log transformed ratio of means (Hedges et al., 1999; Lajeunesse, 2011) for each observation pair (ambient and elevated).
```{r}
df3 <- df2 %>%
  mutate(id = 1:n()) %>% 
  filter(sd_c > 0, sd_t > 0) %>% 
  
  ## keep only essential variables and drop rows containing missing values for essential variables
  select(id, exp = citation, response, nfert_lev, x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>% 
  drop_na() %>% 
  
  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = x_t, sd1i = sd_t, n1i = rep_t, 
    m2i = x_c, sd2i = sd_c, n2i = rep_c, 
    data = ., 
    append = TRUE, 
    var.names = c("logr", "logr_var") 
    ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  ## get standard error
  mutate( logr_se = sqrt(logr_var) / sqrt(rep_t) )
```

## Aggregate by experiment

Aggregate all measurements (species, multiple years, sampling dates and plots) by experiment (and response variable - although here only one) for meta analysis.

### All pooled 

```{r}
df4 <- df3 %>% 
  
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  
  # re-create ID (common ID per experiment and response variable)
  select(-id) %>%
  mutate( id = paste(exp, response, sep = "_XXX_")) %>% 
  
  MAd::agg( 
    id = id, 
    es = logr, 
    var = logr_var,
    cor = 1.0, 
    method = "BHHR", 
    data = . 
    ) %>% 

  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  # separate ID again for ease of data use
  mutate( id = str_split(id, "_XXX_") ) %>%
  mutate( exp = purrr::map_chr(id, 1),
          response = purrr::map_chr(id, 2) ) %>%
  
  ## rename again
  select(exp, response, logr = es, logr_var = var) %>%

  ## add number of observations (sum of plots and repeated samplings)
  left_join(
    df3 %>%
      group_by(exp, response) %>%
      summarise(n_c = sum(rep_c), n_t = sum(rep_t)),
    by = c("exp", "response")
  ) %>% 
  
  ## get standard error. Verify if number available observations are identical
  ## for ambient and elevated. Use N from control here (n_c).
  mutate( logr_se = sqrt(logr_var) / sqrt(n_c) )
```

### By N-fert level

```{r}
df4b <- df3 %>% 
  
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  
  # re-create ID (common ID per experiment and response variable)
  select(-id) %>%
  mutate( id = paste(exp, response, nfert_lev, sep = "_XXX_")) %>% 
  
  MAd::agg( 
    id = id, 
    es = logr, 
    var = logr_var,
    cor = 1.0, 
    method = "BHHR", 
    data = . 
    ) %>% 

  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  # separate ID again for ease of data use
  mutate( id = str_split(id, "_XXX_") ) %>%
  mutate( exp = purrr::map_chr(id, 1),
          response = purrr::map_chr(id, 2),
          nfert_lev = purrr::map_chr(id, 3) ) %>%
  
  ## rename again
  select(exp, response, nfert_lev, logr = es, logr_var = var) %>%

  ## add number of observations (sum of plots and repeated samplings)
  left_join(
    df3 %>%
      group_by(exp, response, nfert_lev) %>%
      summarise(n_c = sum(rep_c), n_t = sum(rep_t)),
    by = c("exp", "response", "nfert_lev")
  ) %>% 
  
  ## get standard error. Verify if number available observations are identical
  ## for ambient and elevated. Use N from control here (n_c).
  mutate( logr_se = sqrt(logr_var) / sqrt(n_c) )
```

## Meta-analysis

### All pooled

Aggregate log-ratios across multiple experiments, taking into account their respective variance and using the experiment identity as a grouping factor for random intercepts.
```{r}
# modify experiment name
get_last <- function(string){
  pos <- str_locate_all(string, "[0-9]") |> 
    map(~max(c(.))) |> 
    unlist()
  return(pos)
}

df4 <- df4 |> 
  mutate(
    loc_last = get_last(exp)
  ) |> 
  mutate(
    exp = str_sub(exp, start = 1, end = loc_last)
  )

out  <- purrr::map(
  as.list(use_vars), 
  ~analyse_meta(df4 %>% 
                  rename(var = response),
                nam_target = .
                ))
names(out) <- use_vars
df_box <- purrr::map_dfr(out, "df_box") |> 
  left_join(
    df4 |> 
      group_by(response) |> 
      summarise(logr_min = min(logr), logr_max = max(logr)) |> 
      rename(var = response),
    by = "var"
  )
saveRDS(df_box, file = "../data/df_box_liang_nfert.rds")
```

### By N-fert level

Separate by fertilisation level
```{r}
df4b <- df4b |>
  filter(response %in% use_vars) |> 
  mutate(var_nfert = paste0(response, "_XXX_", nfert_lev))

use_varsb <- df4b |> 
  pull(var_nfert) |> 
  unique()
  
outb  <- purrr::map(
  as.list(use_varsb), 
  ~analyse_meta(df4b |>  
                  rename(var = var_nfert), 
                nam_target = .
                ))

names(outb) <- use_varsb
df_boxb <- purrr::map_dfr(outb, "df_box") |> 
  separate(var, 
           c("var", "nfert_lev"),
           sep = "_XXX_")
saveRDS(df_boxb, file = "../data/df_boxb_liang_nfert.rds")
```

## Visualisations

### All pooled

Plot dots and my box.
```{r}
mylabs <- c(
  expression(paste(italic("A")[sat])),
  expression(paste(italic("V")[cmax])),
  expression(paste(italic("N")[area])),
  expression(paste(italic("N")[mass])),
  "LAI",
  # "Leaf area",
  "Leaf biomass"
)

df_plot <- df4 |> 
  ## give it a nice order (for plotting)
  filter(response %in% use_vars) |> 
  mutate(response = factor(response, levels = rev(use_vars)))

set.seed(2023)
gg_nfert_liang <- ggplot(
    data = df_plot, 
    aes(x = response, y = logr)) +
  geom_jitter( color = rgb(0,0,0,0.3), 
               aes( size = 1/logr_se ), 
               position = position_jitter(w = 0.2, h = 0), 
               show.legend = FALSE ) +
  geom_crossbar( data = df_box, 
                 aes(x = var, y = middle, ymin = ymin, ymax = ymax), 
                 fill = "royalblue", color = "royalblue4", alpha = 0.6, width = 0.5 ) +
  geom_hline( yintercept = 0.0, size = 0.5, linetype = "dotted") +
  labs(x = "Variable", y = "Log Response Ratio") +
  coord_flip() +
  scale_x_discrete("", labels = rev(mylabs)) +
  ylim(-1, 3.5) +
  labs(x = "", 
       y = "Log response ratio") +
  ggtitle(expression(bold("Response to N")), subtitle = "Liang et al. data") +
  theme_classic() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text( hjust = 0.98, 
                               vjust = 0, 
                               face = 'bold', 
                               margin = margin(t=20, b=-30),
                               size = 17
                               ),
    plot.subtitle = element_text( hjust = 0.98, 
                                  vjust = -3, 
                                  margin = margin(t=20, b=-30),
                                  size = 15
                                  )
  )

saveRDS(gg_nfert_liang, file = paste0(here::here(), "/data/gg_nfert_liang.rds"))

gg_nfert_liang
ggsave(paste0(here::here(), "/fig/metaanalysis_nfert_liang.pdf"), width = 6, height = 4)
ggsave(paste0(here::here(), "/fig/metaanalysis_nfert_liang.png"), width = 6, height = 4)
```

**Findings:**

- Small and non-significant increase in Vcmax, stronger increase in Narea and Nmass.
- Increased stomatal conductance.
- No significant changes in LMA and SLA.
- Strongly positive response in leaf biomass, plant-level leaf area, and ecosystem-level leaf area (LAI).

Below: experiment-level meta analysis plots. Note that variable names written to (console) output are pasted in abvoe the respective "forest plot". 


#### Forest plots

```{r}
for (ivar in unique(use_vars)){
  print(ivar)
  tmp <- out[[ivar]]$modl
  pdf(
    file = here(paste0("fig/forestplot_liangetal_nfert_", ivar, ".pdf")),
    width = 8,
    height = 0.2*nrow(tmp$data) + 4
  )
  try(
    metafor::forest(
      tmp,
      header = c(ivar, "")
      )
    )
  dev.off()
}
# purrr::map(as.list(use_vars), ~try(metafor::forest(out[[.]]$modl)))
```

### By fertilisation level

Plot dots and my box.
```{r}
df_plot <- df4b |> 
  drop_na(response) |> 
  
  ## give it a nice order (for plotting)
  mutate(myvar = factor(response, levels = rev(use_vars)))

gg_nfert_liang_nfert <- ggplot() +
  geom_point( data = df_plot,
               aes(x = myvar, y = logr, size = 1/logr_se, group = nfert_lev),
               alpha = 0.5,
               # position = position_jitter(w = 0.2, h = 0),
               position = position_dodge(0.7),
               show.legend = FALSE
               ) +
  geom_crossbar( data = df_boxb %>% drop_na(var), 
                 aes(x = var, 
                     y = middle, 
                     ymin = ymin, ymax = ymax, 
                     fill = nfert_lev, 
                     group = nfert_lev
                     ),
                 position = position_dodge(0.7),
                 color = "black", 
                 alpha = 0.6,
                 width = 0.5 
                 ) +
  geom_hline( yintercept = 0.0, size = 0.5) +
  coord_flip() +
  # scale_x_discrete("", labels = mylabl) +
  scale_fill_viridis_d(name = "Fertilisation \nlevel", labels = c("low", "mid", "high")) +
  scale_x_discrete("", labels = rev(mylabs)) +
  labs(title = "Response to N-fertilisation",
       subtitle = "Liang et al. 2020 data",
       x = "", 
       y = "Log response ratio")

saveRDS(gg_nfert_liang_nfert, file = paste0(here::here(), "/data/gg_nfert_liang_nfert.rds"))

gg_nfert_liang_nfert
ggsave(paste0(here::here(), "/fig/metaanalysis_nfert_liang_nfert.pdf"), width = 8, height = 6)
ggsave(paste0(here::here(), "/fig/metaanalysis_nfert_liang_nfert.png"), width = 8, height = 6)
```

<!-- # N fertilisation experiments -->

<!-- Select data based on our purpose. -->
<!-- ```{r} -->
<!-- df_sub <- df %>%  -->
<!--   # filter(str_detect(treatment, "f")) %>%  -->
<!--   filter(treatment == "f") %>%  -->
<!--   filter(response %in% c("vcmax", "leaf_n", "leaf_cn")) %>%  -->
<!--   filter(experiment_type %in% c("field", "greenhouse", "growth_chamber", "outdoor_chamber"))  # not "pot" -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_sub %>%  -->
<!--   select(id, exp, response, experiment_type, treatment, x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>%  -->
<!--   drop_na() %>%  -->
<!--   group_by(response) %>%  -->
<!--   summarise(n_data = n()) %>%  -->
<!--   left_join( -->
<!--     df_sub %>%  -->
<!--       select(exp, response) %>%  -->
<!--       drop_na() %>% -->
<!--       distinct() %>%  -->
<!--       group_by(response) %>%  -->
<!--       summarise(n_experiments = n()), -->
<!--     by = "response" -->
<!--   ) -->
<!-- ``` -->


<!-- ## Analysis -->



<!-- ## Comments -->

<!-- - Needs extensive explanation of each column -->
<!-- - No data for Narea or should this be derived from leaf_n and LMA? -->
<!-- - Standardised units per variable (response). Currently, values are given in different units. This could be post-processed quite easily. -->
<!-- - Units of `leaf_cn` is given as `"molar"` or `NA`. Please provide information about whether `NA` means gC/gN. -->
<!-- ```{r} -->
<!-- df %>%  -->
<!--   filter(response %in% c("leaf_n")) %>% -->
<!--   pull(x_units) %>%  -->
<!--   unique() -->

<!-- df %>%  -->
<!--   filter(response %in% c("vcmax")) %>% -->
<!--   pull(x_units) %>%  -->
<!--   unique() -->

<!-- df %>%  -->
<!--   filter(response %in% c("leaf_cn")) %>% -->
<!--   pull(x_units) -->
<!-- ``` -->
