---
title: "Response to N-fert., using combined NutNet + MESI data"
author: "Evan Perkowski and Beni Stocker"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages and data

```{r, warning = FALSE, message=FALSE}
## Libraries
library(dplyr)       # for working with ease
library(tidyr)       # for working with ease
library(stringr)
library(ggplot2)
library(readr)
library(metafor)     # for meta analysis
library(MAd)         # for meta analysis

## MESI data
## IMPORTANT: LOAD DATA FROM REPO AT COMMIT d73bbdb12f469e219e3bb360f4b79fdb0dfa0891 Mon Nov 28 11:37:49 2022 +0100
df_mesi <- read_csv("~/mesi-db_test/data/mesi_main.csv")

## NutNet data from Cleland et al. (2019). Calculate aboveground biomass from root biomass and root mass fraction, then calculate root:shoot ratio
df_nutnet <- read_csv(paste0(here::here(), "/data/Cleland_2019_data.csv")) %>%
  rename(bgb = Rootsgperm2, rmf = rootmassfraction) %>%
  mutate(agb = (bgb/rmf) - bgb,
         root_shoot_ratio = bgb / agb)
```

# NutNet data from Cleland et al. (2019)

Clean data and calculate log RR for root mass fraction, belowground biomass, aboveground biomass, and root:shoot ratio for each site in a single step

```{r, message=FALSE}
df_nutnet.summary <- df_nutnet %>%
  filter(treatment_name == "N" | treatment_name == "Control") %>%
  group_by(site_code, treatment_name) %>%
  summarize(
    
    aridity = mean(Aridity, na.rm = TRUE),
    light = mean(Light, na.rm = TRUE),
    
    x_rmf = mean(rmf, na.rm = TRUE),
    sd_rmf = sd(rmf, na.rm = TRUE),
    n_rmf = length(rmf),
    
    x_bgb = mean(bgb, na.rm = TRUE),
    sd_bgb = sd(bgb, na.rm = TRUE),
    n_bgb = length(bgb),
    
    x_agb = mean(agb, na.rm = TRUE),
    sd_agb = sd(agb, na.rm = TRUE),
    n_agb = length(agb),
            
    x_rootshoot = mean(root_shoot_ratio, na.rm = TRUE),
    sd_rootshoot = sd(root_shoot_ratio, na.rm = TRUE),
    n_rootshoot = length(root_shoot_ratio)
    
    ) %>%
  
  pivot_wider(names_from = treatment_name, values_from = x_rmf:n_rootshoot) %>%
  rename(
    
    x_c_rmf = x_rmf_Control, x_t_rmf = x_rmf_N,
    sd_c_rmf = sd_rmf_Control, sd_t_rmf = sd_rmf_N,
    rep_c_rmf = n_rmf_Control, rep_t_rmf = n_rmf_N,
    
    x_c_bgb = x_bgb_Control, x_t_bgb = x_bgb_N,
    sd_c_bgb = sd_bgb_Control, sd_t_bgb = sd_bgb_N,
    rep_c_bgb = n_bgb_Control, rep_t_bgb = n_bgb_N,
         
    x_c_agb = x_agb_Control, x_t_agb = x_agb_N,
    sd_c_agb = sd_agb_Control, sd_t_agb = sd_agb_N,
    rep_c_agb = n_agb_Control, rep_t_agb = n_agb_N,   
         
    x_c_rootshoot = x_rootshoot_Control, x_t_rootshoot = x_rootshoot_N,
    sd_c_rootshoot = sd_rootshoot_Control, sd_t_rootshoot = sd_rootshoot_N,
    rep_c_rootshoot = n_rootshoot_Control, rep_t_rootshoot = n_rootshoot_N
    )
```

Equivalent?
```{r}
df_nutnet_0 <- df_nutnet |> 
  
  ## N-fertilization-only manipulation
  filter(treatment_name == "N" | treatment_name == "Control")
```

```{r}
df_nutnet_2 <- df_nutnet_0 |> 
  
  # make long to conform with MESI
  pivot_longer(cols = c(bgb, rmf, agb, root_shoot_ratio), 
               names_to = "myvar",
               values_to = "x"
               ) |> 
  
  # aggregate across experiments, treatments, and years
  group_by(site_code, treatment_name, myvar) |> 
  summarise(x_mean = mean(x, na.rm = TRUE),
            x_sd = sd(x, na.rm = TRUE),
            rep = n(),
            .groups = 'drop') |> 
  
  # require at least three repetitions - XXX this is different from Evan's implementation
  filter(rep > 2) |> 
  
  # subset
  select(site_code, myvar, treatment_name, x_mean, x_sd, rep) |> 

  # make wide by treatment
  pivot_wider(names_from = treatment_name, values_from = c(x_mean, x_sd, rep)) |> 
  drop_na()
```


Calculate ROM for RMF data:

```{r, message=FALSE}
df_nutnet.rmf <- df_nutnet.summary[1:9] %>%
  rename(x_c = x_c_rmf,
         x_t = x_t_rmf,
         sd_c = sd_c_rmf,
         sd_t = sd_t_rmf,
         rep_c = rep_c_rmf,
         rep_t = rep_t_rmf) %>%
  metafor::escalc(measure = "ROM", 
                  m1i = x_t, sd1i = sd_t, n1i = rep_t, 
                  m2i = x_c, sd2i = sd_c, n2i = rep_c, 
                  data = ., 
                  append = TRUE, 
                  var.names = c("logr", "logr_var")) %>%
  mutate(logr_se = sqrt(logr_var) / sqrt(rep_t),
         myvar = "rmf")
```


Calculate ROM for belowground biomass:

```{r, message=FALSE}
df_nutnet.bgb <- df_nutnet.summary[, c(1:3, 10:15)] %>%
  rename(x_c = x_c_bgb, x_t = x_t_bgb,
         sd_c = sd_c_bgb, sd_t = sd_t_bgb,
         rep_c = rep_c_bgb, rep_t = rep_t_bgb) %>%
  metafor::escalc( 
    measure = "ROM", 
    m1i = x_t, sd1i = sd_t, n1i = rep_t, 
    m2i = x_c, sd2i = sd_c, n2i = rep_c, 
    data = ., 
    append = TRUE, 
    var.names = c("logr", "logr_var")) %>%
  mutate(logr_se = sqrt(logr_var) / sqrt(rep_t),
         myvar = "bgb")
```


Calculate ROM for aboveground biomass data:

```{r, message=FALSE}
df_nutnet.agb <- df_nutnet.summary[c(1:3, 16:21)] %>%
  rename(x_c = x_c_agb,
         x_t = x_t_agb,
         sd_c = sd_c_agb,
         sd_t = sd_t_agb,
         rep_c = rep_c_agb,
         rep_t = rep_t_agb) %>%
  metafor::escalc(measure = "ROM", 
                  m1i = x_t, sd1i = sd_t, n1i = rep_t, 
                  m2i = x_c, sd2i = sd_c, n2i = rep_c, 
                  data = ., 
                  append = TRUE, 
                  var.names = c("logr", "logr_var")) %>%
  mutate(logr_se = sqrt(logr_var) / sqrt(rep_t),
         myvar = "agb")
```


Calculate ROM for root:shoot ratio data:

```{r, message=FALSE}
df_nutnet.rootshoot <- df_nutnet.summary[c(1:3, 22:27)] %>%
  rename(x_c = x_c_rootshoot,
         x_t = x_t_rootshoot,
         sd_c = sd_c_rootshoot,
         sd_t = sd_t_rootshoot,
         rep_c = rep_c_rootshoot,
         rep_t = rep_t_rootshoot) %>%
  metafor::escalc(measure = "ROM", 
                  m1i = x_t, sd1i = sd_t, n1i = rep_t, 
                  m2i = x_c, sd2i = sd_c, n2i = rep_c, 
                  data = ., 
                  append = TRUE, 
                  var.names = c("logr", "logr_var")) %>%
  mutate(logr_se = sqrt(logr_var) / sqrt(rep_t),
         myvar = "root_shoot_ratio")
```

Equivalent? `df_nutnet_3` should be equivalent to `nutnet.total`.
```{r}
df_nutnet_3 <- df_nutnet_2 %>%
  
  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = x_mean_N,       sd1i = x_sd_N,       n1i = rep_N, 
    m2i = x_mean_Control, sd2i = x_sd_Control, n2i = rep_Control, 
    data = ., 
    append = TRUE, 
    var.names = c("logr", "logr_var") 
    ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  ## get standard error
  mutate( logr_se = sqrt(logr_var) / sqrt(rep_N) ) |> 
  
  # make conform
  select(exp = site_code,
         myvar,
         logr,
         logr_var,
         logr_se,
         n_c = rep_Control,
         n_t = rep_N) |> 
  mutate(exp = str_c("nutnet_", exp),
         treatment = "f")
```


Merge lnRR data for all belowground traits, format for easy merge into MESI dataset
```{r, message=FALSE}
nutnet.total <- df_nutnet.bgb %>% 
  full_join(df_nutnet.rmf) %>%
  full_join(df_nutnet.agb) %>% 
  full_join(df_nutnet.rootshoot) %>%
  dplyr::select(exp = site_code, 
                myvar, 
                logr, 
                logr_var, 
                logr_se,
                n_c = rep_c, 
                n_t = rep_t) %>%
  mutate(exp = str_c("nutnet_", exp),
         treatment = "f")
```

## Test

Are Evan's and mine identical? Yes!
```{r}
df_nutnet_3 |> 
  select(exp, myvar, logr_mine = logr) |> 
  left_join(
    nutnet.total |> 
      select(exp, myvar, logr_evan = logr),
    by = c("exp", "myvar")
  ) |> 
  mutate(diff = logr_mine - logr_evan) |>
  pull(diff)
```


# MESI

Follows exact approach from Beni's analysis from Rpubs. Link: https://rpubs.com/stineb/analysis_mesi

## Data subsetting

```{r, message=FALSE}
df0 <- df_mesi %>%
  
  ## fertilisation experiments only
  filter(treatment == "f") |> 
  
  ## N-fertilisation only (no P or K manipulation)
  filter(npk == "_100") |> 
  
  ## only experiments conducted in the field
  filter(experiment_type == "field") |> 
  
  ## exclude data where N-fertilisation levels are very large 
  filter(n_t <= 30)

# ## Data subsetting
# use_exp <- df0 %>% 
#   filter(!is.na(sampling_year)) %>% 
#   group_by(exp) %>% 
#   summarise(nyears = max(sampling_year)) %>% 
#   filter(nyears >= 3) %>% 
#   pull(exp)

# df1 <- df0 %>%
#   filter(exp %in% use_exp)

## do not subset to minimum experiment length
df1 <- df0
```

## Select variables

```{r, warning=FALSE}
use_response <- c("anpp", 
                  "agb", 
                  "bgb", 
                  "fine_root_biomass", 
                  "root_production", 
                  "fine_root_production", 
                  "root_shoot_ratio", 
                  "soil_no3-n", 
                  "soil_nh4-n", 
                  "soil_nh4", 
                  "soil_no3", 
                  "soil_solution_nh4", 
                  "soil_solution_no3", 
                  "root_n_uptake", 
                  "root_nh4_uptake", 
                  "root_no3_uptake"
                  )

df2 <- df1 %>% 
  filter(response %in% use_response) %>% 
  mutate(myvar = response) %>% 
  mutate(myvar = ifelse(myvar %in%  c("agb", 
                                      "agb_coarse"),   
                        "agb", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("bgb", 
                                      "fine_root_biomass", 
                                      "coarse_root_c_stock", 
                                      "bgb_coarse"), 
                        "bgb", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("bgb", 
                                      "fine_root_biomass"), 
                        "bgb", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("root_production", 
                                      "fine_root_production", 
                                      "coarse_root_production"),
                        "root_production", 
                        myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("root_n_uptake", 
                                      "root_nh4_uptake", 
                                      "root_no3_uptake"),
                        "n_uptake", 
                        myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("soil_no3-n", 
                                      "soil_nh4-n", 
                                      "soil_nh4", 
                                      "soil_no3", 
                                      "soil_solution_nh4", 
                                      "soil_solution_no3"), 
                        "n_inorg", 
                        myvar))

use_vars <- unique(df2$myvar)
```

## Analysis

Calculate `"ROM"` - the log transformed ratio of means (Hedges et al., 1999; Lajeunesse, 2011) for each observation pair (ambient and elevated).

```{r, warning=FALSE}
df3 <- df2 %>%
  
  ## keep only essential variables and drop rows containing missing values for 
  ## essential variables
  select(id, duplicate_id, exp, myvar, treatment, sampling_year, 
         x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>% 
  
  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = x_t, sd1i = sd_t, n1i = rep_t, 
    m2i = x_c, sd2i = sd_c, n2i = rep_c, 
    data = ., 
    append = TRUE, 
    var.names = c("logr", "logr_var") 
  ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  ## get standard error
  mutate( logr_se = sqrt(logr_var) / sqrt(rep_t) )
```


```{r}
# Aggregate all measurements (multiple years, sampling dates and plots) by experiment (and response variable - although here only one) for meta analysis.
df4 <- df3 %>% 
  
  # suggested addition by Kevin, email 02.10.2023 10:03
  dplyr::distinct(duplicate_id, x_t, x_c, .keep_all = TRUE) |> 
  
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  
  # re-create ID (common ID per experiment and response variable)
  select(-id) %>%
  mutate( id = paste(exp, myvar, sep = "_XXX_")) %>% 
  
  MAd::agg( 
    id = id, 
    es = logr, 
    var = logr_var,
    cor = 1.0, 
    method = "BHHR", 
    data = . 
  ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  # separate ID again for ease of data use
  mutate( id = str_split(id, "_XXX_") ) %>%
  mutate( exp = purrr::map_chr(id, 1),
          myvar = purrr::map_chr(id, 2) ) %>%
  
  ## rename again
  select(exp, myvar, logr = es, logr_var = var) %>%
  
  ## add number of observations (sum of plots and repeated samplings)
  left_join(
    df3 %>%
      group_by(exp, myvar, treatment) %>%
      summarise(n_c = sum(rep_c), n_t = sum(rep_t)),
    by = c("exp", "myvar")
  ) %>% 
  
  ## get standard error. Verify if number available observations are identical
  ## for ambient and elevated. Use N from control here (n_c).
  mutate( logr_se = sqrt(logr_var) / sqrt(n_c) )
```

# Combine

```{r}
df4_combined <- df4 |> 
  
  # ## Merge with nutnet dataset
  # bind_rows(nutnet.total)

  ## should be equivalent:
  bind_rows(df_nutnet_3)
```

Add root mass fraction to var list (for plots)
```{r, warning=FALSE}
use_vars2 <- c(use_vars, "rmf")
```

# Meta-analysis

Aggregate log-ratios across multiple experiments, taking into account their respective variance and using the experiment identity as a grouping factor for random intercepts.

```{r, warning=FALSE, message=FALSE}
source(paste0(here::here(), "/R/analyse_meta.R"))

out  <- purrr::map(as.list(use_vars2), 
                   ~analyse_meta(df4_combined %>% 
                                   rename(var = myvar), nam_target = .))
names(out) <- use_vars2
df_box <- purrr::map_dfr(out, "df_box") |> 
  left_join(
    df4_combined |> 
      group_by(myvar) |> 
      summarise(logr_min = min(logr), logr_max = max(logr)) |> 
      rename(var = myvar),
    by = "var"
  )
saveRDS(df_box, file = paste0(here::here(), "/data/df_box_mesi_nfert_eperkowski.rds"))
```

## Final data size

Number of data points (plot-level measurements) per variable:

```{r, message=FALSE}
df4_combined %>% 
  group_by(myvar) %>% 
  summarise(n_plots = sum(n_c, na.rm = TRUE), n_exp = n()) %>% 
  rename_("Variable"="myvar", "N observations"="n_plots", "N experiments"="n_exp") %>% 
  knitr::kable()
```

Number of data points (plot-level measurements) per experiment:
```{r, warning=FALSE}
df4_combined %>% 
  group_by(exp) %>% 
  summarise(n_plots = sum(n_c), n_exp = n()) %>% 
  rename_("Experiment"="exp", "N observations"="n_plots", "N experiments"="n_exp") %>% 
  knitr::kable()
```


# Publication figure

## Response ratio by variable

```{r, warning=FALSE}
mylabl = rev(c(
    "ANPP",
    "AGB",
    "Root NPP",
    "BGB",
    "Root:shoot",
    "RMF",
    "N uptake",
    "Inorganic N"
  ))

set.seed(2023)
gg_nfert_mesinutnet <- df4_combined %>%
  
  drop_na(myvar) %>% 
  
  ## give it a nice order (for plotting)
  mutate(myvar = factor(myvar, levels = rev(c("anpp", 
                                              "agb", 
                                              "root_production", 
                                              "bgb", 
                                              "root_shoot_ratio", 
                                              "rmf", 
                                              "n_uptake", 
                                              "n_inorg")))) %>% 
  ggplot( aes(x = myvar, y = logr)) +
  geom_jitter(color = rgb(0,0,0,0.3), 
              aes( size = 1/logr_se ), 
              position = position_jitter(w = 0.2, h = 0),
              show.legend = FALSE) +
  geom_crossbar( data = df_box %>% drop_na(var), 
                 aes(x = var, y = middle, ymin = ymin, ymax = ymax),
                 fill = "royalblue", 
                 color = "royalblue4", 
                 alpha = 0.6, 
                 width = 0.5 ) +
  geom_hline( yintercept = 0.0, size = 0.5, linetype = "dotted" ) +
  scale_x_discrete("", labels = mylabl) +
  labs(x = "", 
       y = "Log response ratio") +
  coord_flip() +
  ylim(-1, 3.5) +
  ggtitle(expression(bold("Response to N")), subtitle = "MESI + NutNet data") +
  theme_classic() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text( hjust = 0.98, 
                               vjust = 0, 
                               face = 'bold', 
                               margin = margin(t=20, b=-30),
                               size = 17
                               ),
    plot.subtitle = element_text( hjust = 0.98, 
                                  vjust = -3, 
                                  margin = margin(t=20, b=-30),
                                  size = 15
                                  )
  )

saveRDS(gg_nfert_mesinutnet, file = paste0(here::here(), "/data/gg_nfert_mesinutnet.rds"))

gg_nfert_mesinutnet
ggsave(paste0(here::here(), "/fig/metaanalysis_nfert_mesinutnet.pdf"), width = 8, height = 6)
ggsave(paste0(here::here(), "/fig/metaanalysis_nfert_mesinutnet.png"), width = 8, height = 6)
```

## Couplings

Make a wide table with ANPP and BNPP:

```{r, warning=FALSE}
df4_agb_bnpp <- df4_combined %>% 
  filter(myvar %in% c("agb", "bgb")) %>% 
  select(exp, myvar, logr) %>%
  tidyr::spread( myvar, logr )

## add standard error 
df4_agb_bnpp <- df4_combined %>% 
  filter(myvar %in% c("agb", "bgb")) %>% 
  select(exp, 
         myvar, 
         logr_se) %>%
  tidyr::spread( myvar, 
                 logr_se ) %>% 
  rename(se_agb = agb, 
         se_bgb = bgb) %>%
  right_join(
    df4_agb_bnpp, 
    by = "exp"
    ) %>% 
  mutate(se = se_agb * se_bgb)

gg_nfert_cpl_agb_bgb_mesinutnet <- df4_agb_bnpp %>% 
  ggplot(aes(x = bgb, 
             y = agb, 
             label = exp)) +
  geom_point(aes(size =  1/se),
             fill = "royalblue",
             color = "royalblue4",
             alpha = 0.6,
             show.legend = FALSE,
             shape = 21) +
  xlim(-1, 1.5) +
  ylim(-1, 1.5) +
  geom_abline(linetype = "dotted") + 
  ggrepel::geom_text_repel(size = 3, 
                           point.padding = 0.5, 
                           segment.alpha = 0, 
                           color = "grey50"
                           ) +
  labs(x = "Log RR of BGB", 
       y = "Log RR of AGB",
       title = "Response to N",
       subtitle = "MESI + NutNet data") +
  theme_classic()

saveRDS(gg_nfert_cpl_agb_bgb_mesinutnet, 
        file = paste0(here::here(), "/data/gg_nfert_cpl_agb_bgb_mesinutnet.rds"))

gg_nfert_cpl_agb_bgb_mesinutnet
ggsave(paste0(here::here(), "/fig/nfert_cpl_agb_bgb_mesinutnet.pdf"), width = 6, height = 6)
ggsave(paste0(here::here(), "/fig/nfert_cpl_agb_bgb_mesinutnet.png"), width = 6, height = 6)
```
