---
title: "Leaf photosynthetic traits and N from field observations"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rbeni)
library(patchwork)
library(skimr)
library(visreg)
library(ingestr)
library(readr)
```

## Ning Dong's data

The data used here is this leaf photosynthetic traits and N dataset:

*Data reference*

Dong, Ning, Prentice, Iain Colin, Wright, Ian, Wang, Han, Atkin,Owen, Bloomfield, Keith, Domingues, Tomas, Gleason, Sean, Maire, Vincent, Onoda, Yusuke, Poorter, Hendrik, & Smith, Nicholas. (2022). dataset for paper "Leaf nitrogen from the perspective of optimal plant function" (Version v1) [Data set]. Zenodo. https://doi.org/10.5281/zenodo.6831903

*Paper reference*

Dong, N., Prentice, I. C., Wright, I. J., Wang, H., Atkin, O. K., Bloomfield, K. J., Domingues, T. F., Gleason, S. M., Maire, V., Onoda, Y., Poorter, H., & Smith, N. G. (2022). Leaf nitrogen from the perspective of optimal plant function. Journal of Ecology, 00, 1– 18. https://doi.org/10.1111/1365-2745.13967

```{r}
df <- read_csv("~/data/leafn_vcmax_ning_dong/data_leafn_vcmax_ning_dong.csv") %>% 
  rename(lat = Latitude, lon = longitude)
```

Quick overview of data.
```{r}
skim(df)
```

### Distribution of data

```{r, warning=FALSE}
gg1 <- df %>% 
  ggplot(aes(vcmax25_obs)) +
  geom_density() +
  labs(title = "Vcmax25")

gg2 <- df %>% 
  ggplot(aes(Narea)) +
  geom_density() +
  labs(title = "Narea")

gg3 <- df %>% 
  ggplot(aes(Nmass)) +
  geom_density() +
  labs(title = "Nmass")

gg4 <- df %>% 
  ggplot(aes(LMA)) +
  geom_density() +
  labs(title = "LMA")

(gg1 + gg2) /
  (gg3 + gg4)
```

Log-transform all variables to make them closer to normally distributed.

```{r, warning=FALSE}
df <- df %>% 
  mutate(log_vcmax25_obs = log(vcmax25_obs),
         log_Narea = log(Narea),
         log_Nmass = log(Nmass),
         log_LMA = log(LMA),
         )
```

```{r, warning=FALSE}
gg1 <- df %>% 
  ggplot(aes(log_vcmax25_obs)) +
  geom_density() +
  labs(title = "Nmass")

gg2 <- df %>% 
  ggplot(aes(log_Narea)) +
  geom_density() +
  labs(title = "Narea")

gg3 <- df %>% 
  ggplot(aes(log_Nmass)) +
  geom_density() +
  labs(title = "Nmass")

gg4 <- df %>% 
  ggplot(aes(log_LMA)) +
  geom_density() +
  labs(title = "LMA")

(gg1 + gg2) /
  (gg3 + gg4)
```

### Analysis vs soil C:N

This uses the soil C:N data as in Ning's dataset.

```{r, warning=FALSE}
gg1 <- df %>% 
  ggplot(aes(soil_cn, log_vcmax25_obs)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg2 <- df %>% 
  ggplot(aes(soil_cn, log_Narea)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg3 <- df %>% 
  ggplot(aes(soil_cn, log_Nmass)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg4 <- df %>% 
  ggplot(aes(soil_cn, log_LMA)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg1 + gg2 + gg3 + gg4
```


**Issues**

- Why soil C:N in discrete values? Where is it extracted from?

### Alternative soil C:N data extraction

Taking it from ISRIC WISE30sec using ingestr. This also extracts soil pH.

```{r}
filn <- "../data/data_leafn_vcmax_ning_dong_ALTCNPH.csv"
if (!file.exists(filn)){
  
  library(ingestr)
  settings_wise <- get_settings_wise(varnam = c("CNrt", "PHAQ"), layer = 1:7)
  
  df_sites <- df %>% 
    select(sitename = site_id, lon, lat) %>% 
    distinct()
  
  df_wise <- ingest(
    df_sites,
    source    = "wise",
    settings  = settings_wise,
    dir       = "~/data/soil/wise/"
    ) %>% 
    unnest(data)
  
  df <- df %>% 
    left_join(
      df_wise %>% 
        rename(site_id = sitename),
      by = "site_id"
    )
  
  write_csv(df, file = filn)  
  
} else {
  
  df <- read_csv(filn)
  
}
```

Plot with alternative C:N.
```{r, warning=FALSE}
gg1 <- df %>% 
  ggplot(aes(CNrt, log_vcmax25_obs)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg2 <- df %>% 
  ggplot(aes(CNrt, log_Narea)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg3 <- df %>% 
  ggplot(aes(CNrt, log_Nmass)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg4 <- df %>% 
  ggplot(aes(CNrt, log_LMA)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg1 + gg2 + gg3 + gg4
```

### Complement climate variables

Several climatic variables are provided in the dataset. To make analyses at different levels consistent - using the same climate and soil data sources - we re-extract this information here following [this](https://computationales.github.io/ingestr/articles/collect_point_environment.html) *ingestr* workflow, collecting the following variables:

- growth temperature
- growing season mean VPD
- growing season mean PPFD
- N deposition

```{r eval=TRUE}
filn <- "../data/data_leafn_vcmax_ning_dong_ALTCLIM.csv"

ingest_wc_bychunk <- function(idxs){
  
  print(idxs)
  
  ## collect WC data for idxs sites
  out <- ingest(
    df_sites |> 
      slice(idxs),
    source    = "worldclim",
    settings  = settings_wc,
    dir       = "~/data/worldclim/"
    )
  
  ## get growting-season means
  out <- out |> 
  
    unnest(data) |> 
  
    ## add latitude
    left_join(df_sites, by = "sitename") |> 
    
    ## vapour pressure kPa -> Pa
    mutate(vapr = vapr * 1e3) |>
    
    ## PPFD from solar radiation: kJ m-2 day-1 -> mol m−2 s−1 PAR
    mutate(ppfd = 1e3 * srad * kfFEC * 1.0e-6 / (60 * 60 * 24)) |>
  
    ## calculate VPD (Pa) based on tmin and tmax
    rowwise() |> 
    mutate(vpd = ingestr::calc_vpd(eact = vapr, tmin = tmin, tmax = tmax)) |> 
    
    ## calculate growth temperature (average daytime temperature)
    mutate(doy = lubridate::yday(lubridate::ymd("2001-01-15") + months(month - 1))) |> 
    mutate(tgrowth = ingestr::calc_tgrowth(tmin, tmax, lat, doy)) |> 
    
    ## average over growing season (where Tgrowth > 0 deg C)
    group_by(sitename) |> 
    nest() |> 
    mutate(data_growingseason = purrr::map(data, ~get_growingseasonmean(.))) |> 
    unnest(data_growingseason) |> 
    select(-data)
  
  return(out)
}
  
get_growingseasonmean <- function(df){
  df |> 
    filter(tgrowth > 0) |> 
    ungroup() |> 
    summarise(across(c(tgrowth, vpd, ppfd), mean))
}

kfFEC <- 2.04
 
if (!file.exists(filn)){
  
  library(ingestr)

  ## Collect climatic variables from WorldClim----------------------------------  
  df_sites <- df %>% 
    select(sitename = site_id, lon, lat) %>% 
    distinct()

  settings_wc <- list(varnam = c("tmin", "tmax", "vapr", "srad"))

  ## do this in rounds, otherwise too memory demanding
  nchunk <- nrow(df_sites)  # only this extreme variant seems to work
  nrows_chunk <- ceiling(nrow(df_sites)/nchunk)
  irow_chunk <- split(seq(nrow(df_sites)), ceiling(seq_along(seq(nrow(df_sites)))/nrows_chunk))
  
  df_wc <- purrr::map_dfr(
    irow_chunk,
    ~ingest_wc_bychunk(.)
  )
  
  write_csv(df_wc, file = "../data/df_wc.csv")
  
  
  ## Collect N deposition-------------------------------------------------------  
  df_ndep <- ingest(df_sites |> 
                      mutate(sitename = as.character(sitename)) |> 
                      mutate(year_start = 1990, year_end = 2009),
                    source    = "ndep",
                    timescale = "y",
                    dir       = "~/data/ndep_lamarque/",
                    verbose   = FALSE
                    ) |> 
    unnest(cols = data) |> 
    group_by(sitename) |> 
    summarise(noy = mean(noy), nhx = mean(nhx)) |> 
    mutate(ndep = noy + nhx) |> 
    select(-noy, -nhx)
  
  ## Combine data---------------------------------------------------------------
  df <- df %>% 
    left_join(
      df_wc %>% 
        rename(site_id = sitename),
      by = "site_id"
    ) |> 
    left_join(
      df_ndep %>% 
        mutate(site_id = as.numeric(sitename)),
      by = "site_id"
    )
  
  write_csv(df, file = filn)  
  
} else {
  df <- read_csv(filn)
}
```

### Multivariate analysis

#### New predictors

Check distribution of predictor variables.

```{r}
# looks alright
df |> 
  ggplot(aes(tgrowth)) +
  geom_density() +
  labs(title = "Tgrowth")

# vpd needs to be transformed
df |> 
  ggplot(aes(log(vpd))) +
  geom_density() +
  labs(title = "VPD")
df <- df |> 
  mutate(log_vpd = log(vpd))

# ppfd looks ok
df |> 
  ggplot(aes(ppfd)) +
  geom_density() +
  labs(title = "PPFD")

# pH looks ok
df |> 
  ggplot(aes(PHAQ)) +
  geom_density() +
  labs(title = "pH")

# ndep needs to be tranformed
df |> 
  ggplot(aes(log(ndep))) +
  geom_density() +
  labs(title = "N deposition")

df <- df |> 
  mutate(log_ndep = log(ndep))
```

Fit multivariate models.
```{r}
# removed: mgdd0 (don't know what it is)
linmod_vcmax <- lm(log_vcmax25_obs ~ log_vpd  + ppfd + alpha + tgrowth + PHAQ + CNrt + elv + log_ndep, data = df)
linmod_narea <- lm(log_Narea ~       log_vpd  + ppfd + alpha + tgrowth + PHAQ + CNrt + elv + log_ndep, data = df)
linmod_nmass <- lm(log_Nmass ~       log_vpd  + ppfd + alpha + tgrowth + PHAQ + CNrt + elv + log_ndep, data = df)
linmod_lma   <- lm(log_LMA ~         log_vpd  + ppfd + alpha + tgrowth + PHAQ + CNrt + elv + log_ndep, data = df)
```

Partial relationship analysis with soil C:N.
```{r}
gg_vcmax <- visreg(linmod_vcmax, "CNrt", gg = TRUE)
gg_narea <- visreg(linmod_narea, "CNrt", gg = TRUE)
gg_nmass <- visreg(linmod_nmass, "CNrt", gg = TRUE)
gg_lma   <- visreg(linmod_lma,   "CNrt", gg = TRUE)

gg_vcmax <- gg_vcmax +
  theme_classic()
gg_narea <- gg_narea +
  theme_classic()
gg_nmass <- gg_nmass +
  theme_classic()
gg_lma <- gg_lma +
  theme_classic()

(gg_vcmax + gg_narea) /
  (gg_nmass + gg_lma)
```

Partial relationship analysis with N deposition
```{r}
gg_vcmax <- visreg(linmod_vcmax, "log_ndep", gg = TRUE)
gg_narea <- visreg(linmod_narea, "log_ndep", gg = TRUE)
gg_nmass <- visreg(linmod_nmass, "log_ndep", gg = TRUE)
gg_lma   <- visreg(linmod_lma,   "log_ndep", gg = TRUE)

gg_vcmax <- gg_vcmax +
  theme_classic()
gg_narea <- gg_narea +
  theme_classic()
gg_nmass <- gg_nmass +
  theme_classic()
gg_lma <- gg_lma +
  theme_classic()

(gg_vcmax + gg_narea) /
  (gg_nmass + gg_lma)
```



#### Old predictors

```{r}
# removed: mgdd0 (don't know what it is)
linmod_vcmax <- lm(log_vcmax25_obs ~ lnD  + lnppfd + alpha + gdday + soil_ph + CNrt + elv, data = df)
linmod_narea <- lm(log_Narea ~       lnD  + lnppfd + alpha + gdday + soil_ph + CNrt + elv, data = df)
linmod_nmass <- lm(log_Nmass ~       lnD  + lnppfd + alpha + gdday + soil_ph + CNrt + elv, data = df)
linmod_lma   <- lm(log_LMA ~         lnD  + lnppfd + alpha + gdday + soil_ph + CNrt + elv, data = df)

library(visreg)
gg_vcmax <- visreg(linmod_vcmax, "CNrt", gg = TRUE)
gg_narea <- visreg(linmod_narea, "CNrt", gg = TRUE)
gg_nmass <- visreg(linmod_nmass, "CNrt", gg = TRUE)
gg_lma   <- visreg(linmod_lma,   "CNrt", gg = TRUE)

gg_vcmax <- gg_vcmax +
  theme_classic()
gg_narea <- gg_narea +
  theme_classic()
gg_nmass <- gg_nmass +
  theme_classic()
gg_lma <- gg_lma +
  theme_classic()

(gg_vcmax + gg_narea) /
  (gg_nmass + gg_lma)
```

**Findings:**

- No decline in Vcmax with increasing soil C:N. 
- Muted response in Narea.
- Clear decline in Nmass with soil C:N
- Clear increase in LMA with soil C:N

**Issues:**

- Why no MAT-like variable available? Or am I missing something?

### Coefficients

Consider leaf Narea as consisting of a metabolic $N_v$ and a structural $N_s$ leaf N component:
$$
N_\text{area} = N_v + N_s
$$
and that $N_v$ is proportional to $V_\text{cmax25}$, while $N_s$ is proportional to LMA (with a zero-y-axis intercept).
```{r}
linmod_leafn = lm(Narea ~ vcmax25_obs + LMA + 0, data = df)
summary(linmod_leafn)
```
These coefficients can now be used in the CN-model for modelling leaf C:N with prescribed species-specific LMA. 

For reference, the following values are returned:
```{r, eval=FALSE}
# Coefficients:
#              Estimate Std. Error t value Pr(>|t|)    
# vcmax25_obs 0.0136818  0.0004412   31.01   <2e-16 ***
# LMA         0.0109072  0.0001553   70.22   <2e-16 ***
```

The coefficient for vcmax25_obs is in units of gN m-2 / (micro-mol m-2 s-1).

The coefficient for LMA is in units of gN m-2 / (gDM m-2).

Visualise model fit.
```{r}
df <- df |> 
  mutate(narea_pred = predict(linmod_leafn, newdata = df))

out <- df |> 
  analyse_modobs2(mod = "narea_pred", obs = "Narea") 

out$gg +
  xlim(0, 18) + ylim(0, 18)
```

## Ellsworth data

TO EXPLORE. BUT LOW PRIORITY.

The data used here is this leaf photosynthetic traits and N dataset:

*Data reference*

Ellsworth, David; Wright, Ian; Crous, Kristine Y.; Goll, Daniel S; Zaehle, Sönke; Cernusak, Lucas A.; et al. (2022): Convergence in phosphorus constraints to photosynthesis dataset. figshare. Dataset. https://doi.org/10.6084/m9.figshare.20010485.v1 

*Paper reference*

Ellsworth, D.S., Crous, K.Y., De Kauwe, M.G. et al. Convergence in phosphorus constraints to photosynthesis in forests around the world. Nat Commun 13, 5005 (2022). https://doi.org/10.1038/s41467-022-32545-0

```{r}
df_ellsworth <- read_csv("~/data/leafnp_vcmax_ellsworth/Ellsworth_NCOMMS_Figure1and2_fulldata.csv")
skim(df_ellsworth)
```

**Issues**

- No longitude and latitude information in this data. Beni will contact David Ellsworth to ask for the long and lat data.

## TROBIT data

```{r}
df_trobit <- read_csv("~/data/leaf_traits/leaf_soil.csv")
```

```{r}
## 81 sites like this:
df_sites_trobit <- df_trobit %>% 
  select(site, lon, lat, lon_2, lat_2, z, species, Vcmax25, narea, lma, Jmax25, nmass) %>% 
  distinct(site, .keep_all = TRUE)
nrow(df_sites_trobit)

## 68 sites when using lon_2 and lat_2:
df_sites_trobit_test <- df_trobit %>% 
  select(site, lon, lat, lon_2, lat_2, z, species, Vcmax25, narea, lma, Jmax25, nmass) %>% 
  distinct(lon_2, lat_2, .keep_all = TRUE)
nrow(df_sites_trobit_test)

## 89 sites when using lon_2, lat_2, and z (elevation):
df_sites_trobit_test2 <- df_trobit %>% 
  select(site, lon, lat, lon_2, lat_2, z, species, Vcmax25, narea, lma, Jmax25, nmass) %>% 
  distinct(lon_2, lat_2, z, .keep_all = TRUE)
nrow(df_sites_trobit_test2)
```

**Issues**

- When using rounded lon and lat to identify sites, their distinct elevation is not considered, When using `site` to identify site ID, there may still be entries taken from distinct elevations. But in general, elevational differences are small. Therefore use `site` for aggregation.

```{r}
duplicated_sites <- df_sites_trobit_test2 %>% 
  group_by(site) %>% 
  summarise(n=n()) %>% 
  filter(n>1) %>% 
  pull(site)

## note differences in elevation (column z)
df_sites_trobit_test2 %>% 
  filter(site %in% duplicated_sites)
```
Aggregate by site (mean) and take log of variables.
```{r}
df_trobit_sitemean <- df_trobit %>% 
  group_by(site) %>% 
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))

# log-transform
df_trobit_sitemean <- df_trobit_sitemean %>% 
  mutate(across(
    c("Vcmax25", "Jmax25", "narea", "nmass", "leafCN", "lma"), 
    .fns = log, 
    .names = "log_{.col}")) %>% 
  rename(sitename = site) # for later use in ingestr
```

### Analysis vs soil C:N

This uses observed soil C:N data. Do a simple regression (plotting against soil C:N), not controlling for other factors.

```{r, warning=FALSE}
gg1 <- df_trobit_sitemean %>% 
  ggplot(aes(CN, log_Vcmax25)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg2 <- df_trobit_sitemean %>% 
  ggplot(aes(CN, log_narea)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg3 <- df_trobit_sitemean %>% 
  ggplot(aes(CN, log_nmass)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg4 <- df_trobit_sitemean %>% 
  ggplot(aes(CN, log_leafCN)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg5 <- df_trobit_sitemean %>% 
  ggplot(aes(CN, log_lma)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg1 + gg2 + gg3 + gg4 + gg5
```


### Complement climate variables

Several climatic variables are provided in the dataset. To make analyses at different levels consistent - using the same climate and soil data sources - we re-extract this information here following [this](https://computationales.github.io/ingestr/articles/collect_point_environment.html) *ingestr* workflow, collecting the following variables:

- growth temperature
- growing season mean VPD
- growing season mean PPFD
- N deposition

```{r}
filn <- "../data/data_leafn_vcmax_trobit_ALTCLIM.csv"

get_growingseasonmean <- function(df){
  df |> 
    filter(tgrowth > 0) |> 
    ungroup() |> 
    summarise(across(c(tgrowth, vpd, ppfd), mean))
}

kfFEC <- 2.04
 
if (!file.exists(filn)){
  
  ## Collect climatic variables from WorldClim----------------------------------  
  df_sites <- df_trobit_sitemean %>% 
    select(sitename, lon, lat, elv = z) %>% 
    distinct() |> 
    
    ## at least one sitename is missing!
    drop_na(sitename)

  settings_wc <- list(varnam = c("tmin", "tmax", "vapr", "srad"))

  ## collect WC data for idxs sites
  df_wc <- ingest(
    df_sites,
    source    = "worldclim",
    settings  = settings_wc,
    dir       = "~/data/worldclim/"
    )
  
  ## get growting-season means
  df_wc <- df_wc |> 
  
    unnest(data) |> 
  
    ## add latitude
    left_join(df_sites, by = "sitename") |> 
    
    ## vapour pressure kPa -> Pa
    mutate(vapr = vapr * 1e3) |>
    
    ## PPFD from solar radiation: kJ m-2 day-1 -> mol m−2 s−1 PAR
    mutate(ppfd = 1e3 * srad * kfFEC * 1.0e-6 / (60 * 60 * 24)) |>
  
    ## calculate VPD (Pa) based on tmin and tmax
    rowwise() |> 
    mutate(vpd = ingestr::calc_vpd(eact = vapr, tmin = tmin, tmax = tmax)) |> 
    
    ## calculate growth temperature (average daytime temperature)
    mutate(doy = lubridate::yday(lubridate::ymd("2001-01-15") + months(month - 1))) |> 
    mutate(tgrowth = ingestr::calc_tgrowth(tmin, tmax, lat, doy)) |> 
    
    ## average over growing season (where Tgrowth > 0 deg C)
    group_by(sitename) |> 
    nest() |> 
    mutate(data_growingseason = purrr::map(data, ~get_growingseasonmean(.))) |> 
    unnest(data_growingseason) |> 
    select(-data)  
  
  ## Collect N deposition-------------------------------------------------------  
  df_ndep <- ingest(
                    df_sites |> 
                      mutate(year_start = 1990, year_end = 2009),
                    source    = "ndep",
                    timescale = "y",
                    dir       = "~/data/ndep_lamarque/",
                    verbose   = FALSE
                    ) |> 
    unnest(cols = data) |> 
    group_by(sitename) |> 
    summarise(noy = mean(noy), nhx = mean(nhx)) |> 
    mutate(ndep = noy + nhx) |> 
    select(-noy, -nhx)
  
  ## Combine data---------------------------------------------------------------
  df_trobit_sitemean <- df_trobit_sitemean %>% 
    left_join(
      df_wc,
      by = "sitename"
    ) |> 
    left_join(
      df_ndep,
      by = "sitename"
    )
  
  write_csv(df_trobit_sitemean, file = filn)  
  
} else {
  df_trobit_sitemean <- read_csv(filn)
}
```


### Multivariate analysis

Missing: alpha
```{r}
df_trobit_sitemean <- df_trobit_sitemean |> 
  mutate(log_vpd = log(vpd),
         log_ndep = log(ndep))

# removed: mgdd0 (don't know what it is)
linmod_vcmax <- lm(log_Vcmax25 ~     log_vpd  + ppfd + tgrowth + pH + CN + z + log_ndep, 
                   data = df_trobit_sitemean)
linmod_narea <- lm(log_narea ~       log_vpd  + ppfd + tgrowth + pH + CN + z + log_ndep, 
                   data = df_trobit_sitemean)
linmod_nmass <- lm(log_nmass ~       log_vpd  + ppfd + tgrowth + pH + CN + z + log_ndep, 
                   data = df_trobit_sitemean)
linmod_lma   <- lm(log_lma ~         log_vpd  + ppfd + tgrowth + pH + CN + z + log_ndep, 
                   data = df_trobit_sitemean)
linmod_cn   <- lm(log_leafCN ~       log_vpd  + ppfd + tgrowth + pH + CN + z + log_ndep, 
                   data = df_trobit_sitemean)
```

Partial relationship analysis with soil C:N.
```{r}
gg_vcmax <- visreg(linmod_vcmax, "CN", gg = TRUE)
gg_narea <- visreg(linmod_narea, "CN", gg = TRUE)
gg_nmass <- visreg(linmod_nmass, "CN", gg = TRUE)
gg_lma   <- visreg(linmod_lma,   "CN", gg = TRUE)
gg_cn   <- visreg(linmod_cn,     "CN", gg = TRUE)

gg_vcmax <- gg_vcmax +
  theme_classic()
gg_narea <- gg_narea +
  theme_classic()
gg_nmass <- gg_nmass +
  theme_classic()
gg_lma <- gg_lma +
  theme_classic()
gg_cn <- gg_cn +
  theme_classic()

(gg_vcmax + gg_narea) /
  (gg_nmass + gg_lma) / 
  (gg_cn + gg_cn)
```

Partial relationship analysis with N deposition.
```{r}
gg_vcmax <- visreg(linmod_vcmax, "log_ndep", gg = TRUE)
gg_narea <- visreg(linmod_narea, "log_ndep", gg = TRUE)
gg_nmass <- visreg(linmod_nmass, "log_ndep", gg = TRUE)
gg_lma   <- visreg(linmod_lma,   "log_ndep", gg = TRUE)
gg_cn   <- visreg(linmod_cn,     "log_ndep", gg = TRUE)

gg_vcmax <- gg_vcmax +
  theme_classic()
gg_narea <- gg_narea +
  theme_classic()
gg_nmass <- gg_nmass +
  theme_classic()
gg_lma <- gg_lma +
  theme_classic()
gg_cn <- gg_cn +
  theme_classic()

(gg_vcmax + gg_narea) /
  (gg_nmass + gg_lma) / 
  (gg_cn + gg_cn)
```

## C:N by PFT

*Paper reference*

Kattge, J., Bönisch, G., Díaz, S., Lavorel, S., Prentice, I. C., Leadley, P., . . . Wirth, C. (2020). TRY plant trait database – enhanced coverage and open access. Global Change Biology, 26(1), 119-188. doi:10.1111/gcb.14904

```{r}
df <- read_csv("~/scratch/try_kattge_2020/data_cn_pdf.csv")
```

```{r}
df |> 
  ggplot(aes(PFTs, CN)) +
  geom_boxplot(fill = "azure3") +
  coord_flip() +
  theme_classic() +
  labs(x = "PFT", y = expression(paste("C:N"[leaf])))
```