---
title: "Experiments data analysis with Van Groenigen et al., 2014 data"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    #theme: paper
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

```{r, warning=FALSE}
library(tidyverse)   # for working with ease
library(metafor)     # for meta analysis
library(MAd)         # for meta analysis
```

Load data files. 

```{r}
df <- read_csv("~/data/gcme/data_received_190325/soilc_decomp_vangroenigen14.csv")
```

Make data (more) conform with MESI.
```{r}
df1 <- df %>% 
  ## Back-calculate SD of k from column wVk = 1/(SD_ambient^2/k_ambient^2 + SD_elevated^2/k_elevated^2)
  ## assuming that ambient_Sd = elevated_Sd
  rowwise() %>% 
  rename( rep_c = n_ambient, rep_t = n_elevated, x_c = k_ambient, x_t = k_elevated,
           exp = exp_nam_vangroenigen ) %>% 
  mutate( sd_c = x_c * x_t / sqrt(wVk * (x_c^2 + x_t^2)) ) %>%
  mutate( sd_t = sd_c,
          response = "kdecay_soil" )
```

## Data overview

**Issues**

- ...

## Data subsetting

Let's use just data from CO2-only experiments.
```{r}
df2 <- df1 %>% 
  
  ## Use CO2-only treatment
  filter(treatment == "c")
```

## Analysis

Calculate `"ROM"` - the log transformed ratio of means (Hedges et al., 1999; Lajeunesse, 2011) for each observation pair (ambient and elevated).
```{r}
df3 <- df2 %>%
  mutate(id = 1:n()) %>% 
  filter(sd_c > 0, sd_t > 0) %>% 
  
  ## keep only essential variables and drop rows containing missing values for essential variables
  select(id, exp, response, x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>% 
  drop_na() %>% 
  
  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = x_t, sd1i = sd_t, n1i = rep_t, 
    m2i = x_c, sd2i = sd_c, n2i = rep_c, 
    data = ., 
    append = TRUE, 
    var.names = c("logr", "logr_var") 
    ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  ## get standard error
  mutate( logr_se = sqrt(logr_var) / sqrt(rep_t) )
```

Aggregate all measurements (species, multiple years, sampling dates and plots) by experiment (and response variable - although here only one) for meta analysis.
```{r}
df4 <- df3 %>% 
  
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  
  # re-create ID (common ID per experiment and response variable)
  mutate( id = exp ) %>% 
  
  MAd::agg( 
    id = id, 
    es = logr, 
    var = logr_var,
    cor = 1.0, 
    method = "BHHR", 
    data = . 
    ) %>% 

  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  ## rename again
  mutate(response = "kdecay_soil") %>% 
  select(exp = id, response, logr = es, logr_var = var) %>%

  ## add number of observations (sum of plots and repeated samplings)
  left_join(
    df3 %>%
      group_by(exp, response) %>%
      summarise(n_c = sum(rep_c), n_t = sum(rep_t)),
    by = c("exp", "response")
  ) %>% 
  
  ## get standard error. Verify if number available observations are identical
  ## for ambient and elevated. Use N from control here (n_c).
  mutate( logr_se = sqrt(logr_var) / sqrt(n_c) )
```

Aggregate log-ratios across multiple experiments, taking into account their respective variance and using the experiment identity as a grouping factor for random intercepts.
```{r}
source("~/lt_cn_review/R/analyse_meta.R")
out  <- analyse_meta(df4 %>% rename(var = response), nam_target = "kdecay_soil")
df_box <- out$df_box
```


## Visualisations

Plot dots and my box.
```{r}
df4 %>%
  ggplot( aes(x = response, y = logr)) +
  geom_jitter( color = rgb(0,0,0,0.3), aes( size = 1/logr_se ), position = position_jitter(w = 0.2, h = 0) ) +
  geom_crossbar( data = df_box, aes(x = var, y = middle, ymin = ymin, ymax = ymax), fill = "tomato", color = "tomato4", alpha = 0.6, width = 0.5 ) +
  geom_hline( yintercept = 0.0, size = 0.5, linetype = "dotted") +
  labs(x = "Variable", y = "Log Response Ratio", size = expression(paste("Error"^{-1}))) +
  coord_flip() +
  ylim(-1, 1)  +
  labs(title = "Van Groenigen et al. 2014 data", subtitle = "Response to CO2")
```
**Findings:**

- Small and non-significant increase in Vcmax, stronger increase in Narea and Nmass.
- Increased stomatal conductance.
- No significant changes in LMA and SLA.
- Strongly positive response in leaf biomass, plant-level leaf area, and ecosystem-level leaf area (LAI).

Below: experiment-level meta analysis plots. Note that variable names written to (console) output are pasted in abvoe the respective "forest plot". 

```{r}
# Vcmax
metafor::forest(out$Vcmax$modl)

# Narea
metafor::forest(out$Narea$modl)

# Nmass
metafor::forest(out$Nmass$modl)

# LMA
metafor::forest(out$LMA$modl)

# SLA
metafor::forest(out$SLA$modl)
```


<!-- # N fertilisation experiments -->

<!-- Select data based on our purpose. -->
<!-- ```{r} -->
<!-- df_sub <- df %>%  -->
<!--   # filter(str_detect(treatment, "f")) %>%  -->
<!--   filter(treatment == "f") %>%  -->
<!--   filter(response %in% c("vcmax", "leaf_n", "leaf_cn")) %>%  -->
<!--   filter(experiment_type %in% c("field", "greenhouse", "growth_chamber", "outdoor_chamber"))  # not "pot" -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_sub %>%  -->
<!--   select(id, exp, response, experiment_type, treatment, x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>%  -->
<!--   drop_na() %>%  -->
<!--   group_by(response) %>%  -->
<!--   summarise(n_data = n()) %>%  -->
<!--   left_join( -->
<!--     df_sub %>%  -->
<!--       select(exp, response) %>%  -->
<!--       drop_na() %>% -->
<!--       distinct() %>%  -->
<!--       group_by(response) %>%  -->
<!--       summarise(n_experiments = n()), -->
<!--     by = "response" -->
<!--   ) -->
<!-- ``` -->


<!-- ## Analysis -->



<!-- ## Comments -->

<!-- - Needs extensive explanation of each column -->
<!-- - No data for Narea or should this be derived from leaf_n and LMA? -->
<!-- - Standardised units per variable (response). Currently, values are given in different units. This could be post-processed quite easily. -->
<!-- - Units of `leaf_cn` is given as `"molar"` or `NA`. Please provide information about whether `NA` means gC/gN. -->
<!-- ```{r} -->
<!-- df %>%  -->
<!--   filter(response %in% c("leaf_n")) %>% -->
<!--   pull(x_units) %>%  -->
<!--   unique() -->

<!-- df %>%  -->
<!--   filter(response %in% c("vcmax")) %>% -->
<!--   pull(x_units) %>%  -->
<!--   unique() -->

<!-- df %>%  -->
<!--   filter(response %in% c("leaf_cn")) %>% -->
<!--   pull(x_units) -->
<!-- ``` -->
