---
title: "Leaf photosynthetic traits and N from field observations"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rbeni)
library(patchwork)
library(skimr)
```

## Ning Dong's data

The data used here is this leaf photosynthetic traits and N dataset:

*Data reference*

Dong, Ning, Prentice, Iain Colin, Wright, Ian, Wang, Han, Atkin,Owen, Bloomfield, Keith, Domingues, Tomas, Gleason, Sean, Maire, Vincent, Onoda, Yusuke, Poorter, Hendrik, & Smith, Nicholas. (2022). dataset for paper "Leaf nitrogen from the perspective of optimal plant function" (Version v1) [Data set]. Zenodo. https://doi.org/10.5281/zenodo.6831903

*Paper reference*

Dong, N., Prentice, I. C., Wright, I. J., Wang, H., Atkin, O. K., Bloomfield, K. J., Domingues, T. F., Gleason, S. M., Maire, V., Onoda, Y., Poorter, H., & Smith, N. G. (2022). Leaf nitrogen from the perspective of optimal plant function. Journal of Ecology, 00, 1– 18. https://doi.org/10.1111/1365-2745.13967

```{r}
df <- read_csv("~/data/leafn_vcmax_ning_dong/data_leafn_vcmax_ning_dong.csv") %>% 
  rename(lat = Latitude, lon = longitude)
```

Quick overview of data.
```{r}
skim(df)
```

### Distribution of data

```{r, warning=FALSE}
gg1 <- df %>% 
  ggplot(aes(vcmax25_obs)) +
  geom_density() +
  labs(title = "Nmass")

gg2 <- df %>% 
  ggplot(aes(Narea)) +
  geom_density() +
  labs(title = "Narea")

gg3 <- df %>% 
  ggplot(aes(Nmass)) +
  geom_density() +
  labs(title = "Nmass")

gg4 <- df %>% 
  ggplot(aes(LMA)) +
  geom_density() +
  labs(title = "LMA")

(gg1 + gg2) /
  (gg3 + gg4)
```

Log-transform all variables to make them closer to normally distributed.

```{r, warning=FALSE}
df <- df %>% 
  mutate(log_vcmax25_obs = log(vcmax25_obs),
         log_Narea = log(Narea),
         log_Nmass = log(Nmass),
         log_LMA = log(LMA),
         )
```

```{r, warning=FALSE}
gg1 <- df %>% 
  ggplot(aes(log_vcmax25_obs)) +
  geom_density() +
  labs(title = "Nmass")

gg2 <- df %>% 
  ggplot(aes(log_Narea)) +
  geom_density() +
  labs(title = "Narea")

gg3 <- df %>% 
  ggplot(aes(log_Nmass)) +
  geom_density() +
  labs(title = "Nmass")

gg4 <- df %>% 
  ggplot(aes(log_LMA)) +
  geom_density() +
  labs(title = "LMA")

(gg1 + gg2) /
  (gg3 + gg4)
```

### Analysis vs soil C:N

```{r, warning=FALSE}
gg1 <- df %>% 
  ggplot(aes(soil_cn, log_vcmax25_obs)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg2 <- df %>% 
  ggplot(aes(soil_cn, log_Narea)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg3 <- df %>% 
  ggplot(aes(soil_cn, log_Nmass)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg4 <- df %>% 
  ggplot(aes(soil_cn, log_LMA)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg1 + gg2 + gg3 + gg4
```


**Issues**

- Why soil C:N in discrete values? Where is it extracted from?

### Alternative soil C:N data extraction

Taking it from ISRIC WISE30sec

```{r}
filn <- "~/data/leafn_vcmax_ning_dong/data_leafn_vcmax_ning_dong_ALTCN.csv"
if (!file.exists(filn)){
  library(ingestr)
  settings_wise <- get_settings_wise(varnam = c("CNrt"), layer = 1:7)
  
  df_sites <- df %>% 
    select(sitename = site_id, lon, lat) %>% 
    distinct()
  
  df_wise <- ingest(
    df_sites,
    source    = "wise",
    settings  = settings_wise,
    dir       = "~/data/soil/wise/"
    ) %>% 
    unnest(data)
  
  df <- df %>% 
    left_join(
      df_wise %>% 
        rename(site_id = sitename),
      by = "site_id"
    )
  
  write_csv(df, file = filn)  
} else {
  df <- read_csv(filn)
}
```

Plot with alternative C:N.
```{r, warning=FALSE}
gg1 <- df %>% 
  ggplot(aes(CNrt, log_vcmax25_obs)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg2 <- df %>% 
  ggplot(aes(CNrt, log_Narea)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg3 <- df %>% 
  ggplot(aes(CNrt, log_Nmass)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg4 <- df %>% 
  ggplot(aes(CNrt, log_LMA)) +
  geom_point() +
  geom_smooth(color = "red", method = "lm") +
  theme_classic()

gg1 + gg2 + gg3 + gg4
```

### Multivariate analysis

```{r}
linmod_vcmax <- lm(vcmax_obs ~ lnD + mgdd0 + lnppfd + alpha + gdday + soil_ph + CNrt + elv, data = df)
linmod_narea <- lm(Narea ~     lnD + mgdd0 + lnppfd + alpha + gdday + soil_ph + CNrt + elv, data = df)
linmod_nmass <- lm(Nmass ~     lnD + mgdd0 + lnppfd + alpha + gdday + soil_ph + CNrt + elv, data = df)
linmod_lma   <- lm(LMA ~       lnD + mgdd0 + lnppfd + alpha + gdday + soil_ph + CNrt + elv, data = df)

library(visreg)
gg_vcmax <- visreg(linmod_vcmax, "CNrt", gg = TRUE)
gg_narea <- visreg(linmod_narea, "CNrt", gg = TRUE)
gg_nmass <- visreg(linmod_nmass, "CNrt", gg = TRUE)
gg_lma   <- visreg(linmod_lma, "CNrt", gg = TRUE)

gg_vcmax <- gg_vcmax +
  theme_classic()
gg_narea <- gg_narea +
  theme_classic()
gg_nmass <- gg_nmass +
  theme_classic()
gg_lma <- gg_lma +
  theme_classic()

(gg_vcmax + gg_narea) /
  (gg_nmass + gg_lma)
```

### Analysis by PFT

```{r}
# -> Ning, please add your analysis as R code here
```

## Ellsworth data

The data used here is this leaf photosynthetic traits and N dataset:

*Data reference*

Ellsworth, David; Wright, Ian; Crous, Kristine Y.; Goll, Daniel S; Zaehle, Sönke; Cernusak, Lucas A.; et al. (2022): Convergence in phosphorus constraints to photosynthesis dataset. figshare. Dataset. https://doi.org/10.6084/m9.figshare.20010485.v1 

*Paper reference*

Ellsworth, D.S., Crous, K.Y., De Kauwe, M.G. et al. Convergence in phosphorus constraints to photosynthesis in forests around the world. Nat Commun 13, 5005 (2022). https://doi.org/10.1038/s41467-022-32545-0

```{r}
df_ellsworth <- read_csv("~/data/leafnp_vcmax_ellsworth/Ellsworth_NCOMMS_Figure1and2_fulldata.csv")
skim(df_ellsworth)
```

**Issues**

- No longitude and latitude information in this data.

## TROBIT data

-> Yunke, please add your analysis code here
