---
title: "Ballpark benchmarking of cnmodel"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rbeni)
library(patchwork)
library(skimr)
```

# Ballpark benchmarking

## Load model output

Outputs are generated with `rsofun/analysis/example_cnmodel_ch_oe1.R`. Copy them to `./data/`. This is for a simulation for FLUXNET site CH-Oe1, a grassland. 

```{r}
system("cp ~/rsofun/tmp/output_cnmodel_ch_oe1.R ../data/")
df_rsofun <- readRDS("../data/output_cnmodel_ch_oe1.R")
# df_rsofun <- read_csv("../data/output_cnmodel_co2.csv")

df_rsofun_agg <- df_rsofun |> 
  summarise(across(where(is.numeric), mean))

adf_rsofun <- df_rsofun |> 
  as_tibble() |> 
  mutate(year = lubridate::year(date)) |> 
  group_by(year) |> 
  summarise(gpp = sum(gpp)) |> 
  ungroup() |> 
  summarise(gpp = mean(gpp))

doydf_rsofun <- df_rsofun |> 
  as_tibble() |> 
  mutate(doy = lubridate::yday(date)) |> 
  group_by(doy) |> 
  summarise(gpp = mean(gpp), fapar = mean(fapar))
```


## Traits

Plot distribution of all observational values (ridge plot) and cnmodel (big dot at the bottom)

- LMA
- Narea
- Nmass
- Vcmax

Get data from Ning's dataset.
```{r}
df <- read_csv("~/data/leafn_vcmax_ning_dong/data_leafn_vcmax_ning_dong.csv") %>% 
  rename(lat = Latitude, lon = longitude)
```

Plot cnmodel outputs on top of distribution of observations of non-woody vegetation.
```{r}
df2 <- df |> 
  filter(woody == "non-woody") |> 
  rename(lma = LMA, narea = Narea, nmass = Nmass, vcmax = vcmax_obs)

gg1 <- ggplot() +
  geom_density(aes(lma, ..density..), data = df2, fill = "#777055ff", color = NA, alpha = 0.5) +
  theme_classic() +
  geom_point(aes(x = lma, y = 0), data = df_rsofun_agg, size = 5, color = "#29a274ff") +
  labs(x = "LMA") +
  coord_cartesian(clip = 'off') +
  scale_y_continuous(expand = c(0, 0))

gg2 <- ggplot() +
  geom_density(aes(nmass, ..density..), data = df2, fill = "#777055ff", color = NA, alpha = 0.5) +
  theme_classic() +
  geom_point(aes(x = narea/lma, y = 0), data = df_rsofun_agg, size = 5, color = "#29a274ff") +
  labs(x = expression(italic("N")[mass])) +
  coord_cartesian(clip = 'off') +
  scale_y_continuous(expand = c(0, 0))

gg3 <- ggplot() +
  geom_density(aes(narea, ..density..), data = df2, fill = "#777055ff", color = NA, alpha = 0.5) +
  theme_classic() +
  geom_point(aes(x = narea, y = 0), data = df_rsofun_agg, size = 5, color = "#29a274ff") +
  labs(x = expression(italic("N")[area])) +
  coord_cartesian(clip = 'off') +
  scale_y_continuous(expand = c(0, 0))

## Vcmax in micro-mol m-2 s-1
gg4 <- ggplot() +
  geom_density(aes(vcmax, ..density..), data = df2, fill = "#777055ff", color = NA, alpha = 0.5) +
  theme_classic() +
  geom_point(aes(x = 1e6 * vcmax, y = 0), data = df_rsofun_agg, size = 5, color = "#29a274ff") +
  labs(x = expression(italic("V")[cmax])) +
  coord_cartesian(clip = 'off') +
  scale_y_continuous(expand = c(0, 0))

gg1 /
  gg2 / 
  gg3 /
  gg4
```

## Fluxes

### GPP

Plot distribution of annual total in temperate, not water-limited grasslands, obtained from FLUXNET

```{r}
library(ingestr)

settings_fluxnet <- list(
  getswc       = FALSE,
  filter_ntdt  = FALSE,
  threshold_GPP= 0.8,
  remove_neg   = FALSE
  )

adf_fluxnet <- ingest(
  siteinfo = siteinfo_fluxnet2015 |> filter(classid == "GRA" & str_detect(koeppen_code, "^C") ),
  source   = "fluxnet",
  getvars  = list(gpp = "GPP_NT_VUT_REF"),
  dir      = "~/data/FLUXNET-2015_Tier1/20191024/YY/", # adjust this with your local path
  settings = settings_fluxnet,
  timescale= "y"
  )

adf_fluxnet <- adf_fluxnet |> 
  mutate(data = purrr::map(data, ~summarise(., gpp = mean(gpp)))) |> 
  unnest(data)
```

```{r}
ggplot() +
  geom_histogram(data = adf_fluxnet, aes(gpp, ..count..), 
                 fill = "#777055ff", color = NA, alpha = 0.5,
                 binwidth = 200) +
  theme_classic() +
  geom_point(aes(x = gpp, y = 0), data = adf_rsofun, size = 10, color = "#29a274ff") +
  labs(x = expression(paste("GPP (gC m"^-2, "yr"^-1, ")"))) +
  coord_cartesian(clip = 'off') +
  scale_y_continuous(expand = c(0, 0))
```

### NPP

Plot distribution of total annual biomass production of unmanaged grasslands, obtained from Matteo’s dataset.

## Allocation

Plot distribution of root mass fractions in temperate, not water-limited grasslands

# Seasonality benchmarking

→ Master student

## fAPAR

- Plot mean seasonal cycle of observations and modelled fAPAR of different grassland (FLUXNET2015) sites separately.

```{r}
filn <- "../data/ddf_modis_CH-Oe1.rds"
if (!file.exists(filn)){
  
  settings_modis <- get_settings_modis(
    bundle            = "modis_fpar",
    data_path         = tempdir(),
    method_interpol   = "loess",
    network           = "fluxnet",
    keep              = TRUE,
    overwrite_raw     = FALSE,
    overwrite_interpol= TRUE,
    n_focal           = 0
    )
  
  df_modis_fpar <- ingest(
    siteinfo_fluxnet2015 |> filter(sitename == "CH-Oe1"),
    source = "modis",
    settings = settings_modis,
    parallel = FALSE,
    ncores = 1
    )
  
  df_modis_fpar <- df_modis_fpar |> 
    unnest(data) |> 
    rename(fapar = modisvar_filled)

  write_rds(df_modis_fpar, file = filn)
  
} else {
  
  df_modis_fpar <- readRDS(filn)
  
}

doydf_modis <- df_modis_fpar |> 
  unnest(data) |> 
  mutate(doy = lubridate::yday(date)) |> 
  group_by(doy) |> 
  summarise(fapar = mean(fapar, na.rm = TRUE))
```   

```{r}
ggplot() +
  geom_line(data = doydf_modis,  aes(doy, fapar), color = "#777055ff") +
  geom_line(data = doydf_rsofun, aes(doy, fapar), color = "#29a274ff") +
  theme_classic()
```

## Fluxes

### GPP

Plot mean seasonal cycle of observations and modelled GPP and NEE of different grassland (FLUXNET2015) sites separately.
    
```{r}
filn <- "../data/ddf_fluxnet_CH-Oe1.rds"
if (!file.exists(filn)){
  settings_fluxnet <- list(
    getswc       = FALSE,
    filter_ntdt  = FALSE,
    threshold_GPP= 0.8,
    remove_neg   = FALSE
    )
  
  ddf_fluxnet <- ingest(
    siteinfo = siteinfo_fluxnet2015 |> filter(sitename == "CH-Oe1"),
    source   = "fluxnet",
    getvars  = list(gpp = "GPP_NT_VUT_REF"),
    dir      = "~/data/FLUXNET-2015_Tier1/20191024/DD/", # adjust this with your local path
    settings = settings_fluxnet,
    timescale= "d"
    )
  write_rds(ddf_fluxnet, file = filn)
} else {
  ddf_fluxnet <- readRDS(filn)
}

doydf_fluxnet <- ddf_fluxnet |> 
  unnest(data) |> 
  mutate(doy = lubridate::yday(date)) |> 
  group_by(doy) |> 
  summarise(gpp = mean(gpp))
```    


```{r}
ggplot() +
  geom_line(data = doydf_fluxnet, aes(doy, gpp), color = "#777055ff") +
  geom_line(data = doydf_rsofun, aes(doy, gpp), color = "#29a274ff") +
  theme_classic()
```
    
### NEE    
    
- N2O
    - Plot time series from simulation where forcing is well-known (building on what Yunke started)
