---
title: "Comparison of CN-model results againtst MESI"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    #theme: paper
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

```{r, warning=FALSE}
library(dplyr)       # for working with ease
library(tidyr)       # for working with ease
library(stringr)
library(ggplot2)
library(readr)
library(metafor)     # for meta analysis
library(MAd)         # for meta analysis
```

## CO2 experiments

Load CO2 meta-analysis results from MESI (produced with `analysis_mesi.Rmd`). 
```{r}
df_box_co2 <- read_rds("../data/df_box_mesi_co2.rds")
```

Add model outputs to this data frame. This is produced with `rsofun/analysis/example_cnmodel.R`.
```{r}
df_cnmodel_co2 <- read_csv("../data/output_cnmodel_co2.csv")

## aggregate before and after
years_before <- 2007:2009
years_after  <- 2015:2025

df_cnmodel_co2_before <- df_cnmodel_co2 |> 
  mutate(root_shoot_ratio = croot / cleaf,
         leaf_n_mass = narea/lma,
         leaf_cn = cleaf / nleaf) |> 
  mutate(year = lubridate::year(date)) |> 
  filter(year %in% years_before) |> 
  summarise(across(where(is.numeric), mean))

df_cnmodel_co2_after <- df_cnmodel_co2 |> 
  mutate(root_shoot_ratio = croot / cleaf,
         leaf_n_mass = narea/lma,
         leaf_cn = cleaf / nleaf) |> 
  mutate(year = lubridate::year(date)) |> 
  filter(year %in% years_after) |> 
  summarise(across(where(is.numeric), mean))

df_cnmodel_co2_logresponse <- log(df_cnmodel_co2_after / df_cnmodel_co2_before) |> 
  mutate(agb = cleaf) |> 
  rename(n_uptake = nup, agb = cleaf, bgb = croot, leaf_n_area = narea, 
         n_inorg = ninorg, gs = gs_accl, leaf_biomass = cleaf) |> 
  pivot_longer(cols = everything(), names_to = "var", values_to = "logrr_cnmodel")
```


Visualisation
```{r}
df_box_co2 |> 
  left_join(df_cnmodel_co2_logresponse, by = "var") |> 
  ## give it a nice order (for plotting)
  mutate(var = factor(var, levels = rev(c("asat", "gpp", "vcmax", "jmax", "gs", "leaf_n_area", "leaf_n_mass", "leaf_cn", "lai", "leaf_area", "leaf_biomass", "anpp", "agb", "root_production", "bgb", "root_shoot_ratio", "n_uptake", "n_inorg")))) %>% 
  
  ggplot() +
  geom_crossbar(aes(x = var, y = middle, ymin = ymin, ymax = ymax), 
                 fill = "tomato", color = "black", alpha = 0.6, width = 0.5 
                 ) +
  geom_point(aes(x = var, y = logrr_cnmodel), size = 3, color = "royalblue" ) +
  geom_hline( yintercept = 0.0, size = 0.5 ) +
  labs(x = "Variable", y = "Log Response Ratio", size = expression(paste("Error"^{-1}))) +
  coord_flip() +
  # ylim(-1, 1) +
  labs(title = "MESI data", subtitle = "Response to eCO2")
```

## N-fertilisation experiments

Load CO2 meta-analysis results from MESI (produced with `analysis_mesi.Rmd`). 
```{r}
df_box_nfert <- read_rds("../data/df_box_mesi_nfert.rds")
```

Add model outputs to this data frame. This is produced with `rsofun/analysis/example_cnmodel.R`.
```{r}
df_cnmodel_nfert <- read_csv("../data/output_cnmodel_nfert.csv")

## aggregate before and after
years_before <- 2007:2009
years_after  <- 2015:2025

df_cnmodel_nfert_before <- df_cnmodel_nfert |> 
  mutate(root_shoot_ratio = croot / cleaf,
         leaf_n_mass = narea/lma,
         leaf_cn = cleaf / nleaf) |> 
  mutate(year = lubridate::year(date)) |> 
  filter(year %in% years_before) |> 
  summarise(across(where(is.numeric), mean))

df_cnmodel_nfert_after <- df_cnmodel_nfert |> 
  mutate(root_shoot_ratio = croot / cleaf,
         leaf_n_mass = narea/lma,
         leaf_cn = cleaf / nleaf) |> 
  mutate(year = lubridate::year(date)) |> 
  filter(year %in% years_after) |> 
  summarise(across(where(is.numeric), mean))

df_cnmodel_nfert_logresponse <- log(df_cnmodel_nfert_after / df_cnmodel_nfert_before) |> 
  mutate(agb = cleaf) |> 
  rename(n_uptake = nup, agb = cleaf, bgb = croot, leaf_n_area = narea, 
         n_inorg = ninorg, gs = gs_accl, leaf_biomass = cleaf) |> 
  pivot_longer(cols = everything(), names_to = "var", values_to = "logrr_cnmodel")
```


Visualisation
```{r}
df_box_nfert |> 
  left_join(df_cnmodel_nfert_logresponse, by = "var") |> 
  ## give it a nice order (for plotting)
  mutate(var = factor(var, levels = rev(c("anpp", "agb", "root_production", "bgb", "root_shoot_ratio", "n_uptake", "n_inorg")))) %>% 
  
  ggplot() +
  geom_crossbar(aes(x = var, y = middle, ymin = ymin, ymax = ymax), 
                 fill = "tomato", color = "black", alpha = 0.6, width = 0.5 
                 ) +
  geom_point(aes(x = var, y = logrr_cnmodel), size = 3, color = "royalblue" ) +
  geom_hline( yintercept = 0.0, size = 0.5 ) +
  labs(x = "Variable", y = "Log Response Ratio", size = expression(paste("Error"^{-1}))) +
  coord_flip() +
  # ylim(-1, 1) +
  labs(title = "MESI data", subtitle = "Response to eCO2")
```