---
title: "Comparison of CN-model results againtst MESI"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    #theme: paper
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

```{r, warning=FALSE}
library(dplyr)       # for working with ease
library(tidyr)       # for working with ease
library(stringr)
library(ggplot2)
library(readr)
library(metafor)     # for meta analysis
library(MAd)         # for meta analysis
```

## CO2 experiments

Load CO2 meta-analysis results from MESI (produced with `vignettes/analysis_mesi.Rmd`). 
```{r}
df_box_co2 <- read_rds("../data/df_box_mesi_co2.rds")
```

Add model outputs to this data frame. This is produced with `analysis/exp_co2_cnmodel.R`.
```{r}
df_cnmodel_co2 <- read_csv(paste0(here::here(), "/data/output_cnmodel_co2_50c01ecbac0ad20114dc9cc28d67006af45f128e.csv"))

## aggregate before and after
years_before <- 2007:2009
years_after  <- 2015:2025

df_cnmodel_co2_before <- df_cnmodel_co2 |> 
  mutate(root_shoot_ratio = croot / cleaf,
         leaf_n_mass = narea/lma,
         leaf_cn = cleaf / nleaf) |> 
  mutate(year = lubridate::year(date)) |> 
  filter(year %in% years_before) |> 
  summarise(across(where(is.numeric), mean))

df_cnmodel_co2_after <- df_cnmodel_co2 |> 
  mutate(root_shoot_ratio = croot / cleaf,
         leaf_n_mass = narea/lma,
         leaf_cn = cleaf / nleaf) |> 
  mutate(year = lubridate::year(date)) |> 
  filter(year %in% years_after) |> 
  summarise(across(where(is.numeric), mean))

df_cnmodel_co2_logresponse <- df_cnmodel_co2_before  |> 
  mutate(agb = cleaf,
         anpp = npp_leaf,
         root_production = npp_root) |> 
  rename(n_uptake = nup, agb = cleaf, bgb = croot, leaf_n_area = narea, 
         n_inorg = ninorg, gs = gs_accl, leaf_biomass = cleaf) |> 
  pivot_longer(cols = everything(), names_to = "var", values_to = "ambient") |> 
  
  left_join(
    df_cnmodel_co2_after  |> 
      mutate(agb = cleaf,
             anpp = npp_leaf,
             root_production = npp_root) |> 
      rename(n_uptake = nup, agb = cleaf, bgb = croot, leaf_n_area = narea, 
             n_inorg = ninorg, gs = gs_accl, leaf_biomass = cleaf) |> 
      pivot_longer(cols = everything(), names_to = "var", values_to = "elevated"),
    by = "var"
  ) |> 
  
  # response ratio, normalised for the doubling of CO2
  mutate(logrr_cnmodel = log(elevated / ambient) / log(2))

# df_cnmodel_co2_logresponse <- log(df_cnmodel_co2_after / df_cnmodel_co2_before) / log(2)  |> 
#   select(cleaf, npp_leaf, npp_root, nup, croot, narea, ninorg, gs_accl) |> 
#   mutate(agb = cleaf,
#          anpp = npp_leaf,
#          root_production = npp_root) |> 
#   rename(n_uptake = nup, agb = cleaf, bgb = croot, leaf_n_area = narea, 
#          n_inorg = ninorg, gs = gs_accl, leaf_biomass = cleaf) |> 
#   pivot_longer(cols = everything(), names_to = "var", values_to = "logrr_cnmodel")
```

Visualisation

```{r}
var_order <- c("asat", 
               "gpp",
               "vcmax", 
               "jmax",
               "leaf_n_area", 
               "leaf_n_mass", 
               "leaf_cn", 
               "lai", 
               # "leaf_biomass", 
               "anpp", 
               "agb", 
               "root_production",
               "bgb",
               "root_shoot_ratio",
               "n_uptake", 
               "n_inorg" 
               )

# mylabl = rev(c(
#     expression(paste(italic("A")[sat])),
#     "GPP",
#     expression(paste(italic("V")[cmax])),
#     expression(paste(italic("J")[max])),
#     expression(paste("Leaf ", italic("N")[area])),
#     expression(paste("Leaf ", italic("N")[mass])),
#     "Leaf C:N",
#     "LAI",
#     # "Leaf mass",
#     "ANPP",
#     "AGB",
#     "Root NPP",
#     "BGB",
#     "Root:shoot",
#     "N uptake",
#     "Inorganic N"
#   ))

show_vars <- c("asat",
               "gpp", 
               "vcmax", 
               "jmax", 
               "leaf_n_area", 
               "leaf_n_mass", 
               "leaf_cn", 
               "lai", 
               # "leaf_area", 
               # "leaf_biomass", 
               "anpp", "agb", 
               "root_production", 
               "bgb", 
               "root_shoot_ratio", 
               "n_uptake", 
               "n_inorg"
               )

df_labels <- tibble(
  myvar = rev(show_vars),
  mylabl = rev(c(
    expression(paste(italic("A")[sat])),
    "GPP",
    expression(paste(italic("V")[cmax])),
    expression(paste(italic("J")[max])),
    expression(paste(italic("N")[area])),
    expression(paste(italic("N")[mass])),
    "Leaf C:N",
    "LAI",
    # "Leaf area",
    # "Leaf mass",
    "ANPP",
    "AGB",
    "Root NPP",
    "BGB",
    "Root:shoot",
    "N uptake",
    "Inorganic N"
  ))
)

df_plot_co2 <- df_box_co2 |> 
  
  left_join(df_cnmodel_co2_logresponse, by = "var") |> 
  
  ## give it a nice order (for plotting)
  filter(var %in% var_order) |> 
  mutate(var = factor(var, levels = rev(var_order)))

df_correct <- df_plot_co2 |> 
  mutate(correct_ci = ifelse(logrr_cnmodel > ymin & logrr_cnmodel < ymax,
                          TRUE, 
                          FALSE),
         correct_all = ifelse(logrr_cnmodel > logr_min & logrr_cnmodel < logr_max,
                          TRUE, 
                          FALSE),
         type = "co2") |> 
  select(var, type, correct_ci, correct_all)

gg_mesi_cnmodel_eco2 <- df_plot_co2 |>
  
  ggplot() +
  geom_crossbar(aes(x = var, y = middle, ymin = ymin, ymax = ymax), 
                fill = "#777055ff", 
                color = "black", 
                alpha = .5, 
                width = 0.5 
                ) +
  geom_segment(aes(x = var, xend = var, y = ymin, yend = logr_min)) +
  geom_segment(aes(x = var, xend = var, y = ymax, yend = logr_max)) +
  geom_point(aes(x = var, y = logrr_cnmodel), 
             size = 5, 
             shape = 21, 
             color = "black", 
             fill = "#29a274ff" 
             ) +
  geom_hline( yintercept = 0.0, size = 0.5, linetype = "dotted") +
  scale_x_discrete("", labels = df_labels$mylabl) +
  coord_flip() +
  ylim(-1, 3.5) +
  labs(x = "", 
       y = "Normalised log response ratio") +
  scale_x_discrete("", labels = df_labels$mylabl) +
  coord_flip() +
  ggtitle(expression(bold("Response to CO"[2])), subtitle = "CN-model vs. MESI data") +
  theme_classic() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text( hjust = 0.98, 
                               vjust = -5, 
                               face = 'bold', 
                               margin = margin(t=20, b=-30),
                               size = 17
                               ),
    plot.subtitle = element_text( hjust = 0.98, 
                                  vjust = -8, 
                                  margin = margin(t=20, b=-30),
                                  size = 15
                                  )
  )

gg_mesi_cnmodel_eco2
ggsave("../fig/mesi_cnmodel_co2.pdf")
```

## N-fertilisation experiments

Load CO2 meta-analysis results from MESI (produced with `vignettes/analysis_mesi.Rmd`). 
```{r}
#df_box_nfert_mesi <- read_rds(paste0(here::here(), "/data/df_box_mesi_nfert.rds"))
df_box_nfert_mesi <- read_rds(file = paste0(here::here(), "/data/df_box_mesi_nfert_eperkowski.rds"))

# read and change variable names to match with cnmodel output
df_box_nfert_liang <- read_rds(paste0(here::here(), "/data/df_box_liang_nfert.rds")) |> 
  mutate(var = ifelse(var == "Aarea", "asat",
           ifelse(var == "Vcmax", "vcmax",
                      ifelse(var == "Narea", "narea",
                             ifelse(var == "Nmass", "nmass",
                                    ifelse(var == "LAI", "lai",
                                           ifelse(var == "Leaf biomass", "leaf_biomass", var)))))))
```

Get log response ratios from model outputs. This is produced with `analysis/exp_nfert_cnmodel.R`.
```{r}
df_cnmodel_nfert <- read_csv(paste0(here::here(), "/data/output_cnmodel_nfert_50c01ecbac0ad20114dc9cc28d67006af45f128e.csv"))

## aggregate before and after
years_before <- 2007:2009
years_after  <- 2015:2025

df_cnmodel_nfert_before <- df_cnmodel_nfert |> 
  mutate(root_shoot_ratio = croot / cleaf,
         nmass = narea/lma,
         leaf_cn = cleaf / nleaf, 
         rmf = croot / (croot + cleaf)) |> 
  mutate(year = lubridate::year(date)) |> 
  filter(year %in% years_before) |> 
  summarise(across(where(is.numeric), mean))

df_cnmodel_nfert_after <- df_cnmodel_nfert |> 
  mutate(root_shoot_ratio = croot / cleaf,
         nmass = narea/lma,
         leaf_cn = cleaf / nleaf, 
         rmf = croot / (croot + cleaf)) |> 
  mutate(year = lubridate::year(date)) |> 
  filter(year %in% years_after) |> 
  summarise(across(where(is.numeric), mean))

df_cnmodel_nfert_logresponse <- 
  log(df_cnmodel_nfert_after / df_cnmodel_nfert_before) |> 
  mutate(agb = cleaf,
         anpp = npp_leaf,
         root_production = npp_root) |> 
  rename(n_uptake = nup, agb = cleaf, bgb = croot,
         n_inorg = ninorg, gs = gs_accl, leaf_biomass = cleaf) |> 
  pivot_longer(cols = everything(), names_to = "var", values_to = "logrr_cnmodel")
```

Visualisation of variables from MESI + NutNet.
```{r}
var_order <- c("anpp", 
               "agb", 
               "root_production", 
               "bgb", 
               "root_shoot_ratio", 
               "rmf",
               "n_uptake", 
               "n_inorg"
               )

mylabl = rev(c(
    "ANPP",
    "AGB",
    "Root NPP",
    "BGB",
    "Root:shoot",
    "RMF",
    "N uptake",
    "Inorganic N"
  ))

df_plot_nfert <- df_box_nfert_mesi |> 
  left_join(df_cnmodel_nfert_logresponse, by = "var") |> 
  
  ## give it a nice order (for plotting)
  filter(var %in% var_order) |> 
  mutate(var = factor(var, levels = rev(var_order)))

df_correct <- df_correct |> 
  bind_rows(
    df_plot_nfert |> 
      mutate(correct_ci = ifelse(logrr_cnmodel > ymin & logrr_cnmodel < ymax,
                              TRUE, 
                              FALSE),
             correct_all = ifelse(logrr_cnmodel > logr_min & logrr_cnmodel < logr_max,
                              TRUE, 
                              FALSE),
             type = "nfert") |> 
      select(var, type, correct_ci, correct_all)
  )
  
gg_mesi_cnmodel_nfert_mesi <- df_plot_nfert |>  
  
  ggplot() +
  geom_crossbar(aes(x = var, y = middle, ymin = ymin, ymax = ymax), 
                fill = "#777055ff", 
                color = "black", 
                alpha = 0.6,
                width = 0.5 
                ) +
  geom_segment(aes(x = var, xend = var, y = ymin, yend = logr_min)) +
  geom_segment(aes(x = var, xend = var, y = ymax, yend = logr_max)) +
  geom_point(aes(x = var, y = logrr_cnmodel), 
             size = 5, 
             shape = 21, 
             color = "black", 
             fill = "#29a274ff" 
             ) +
  geom_hline( yintercept = 0.0, size = 0.5, linetype = "dotted") +
  coord_flip() +
  scale_x_discrete("", labels = mylabl) +
  ylim(-1, 3.5) +
  labs(x = "", 
       y = "Log response ratio") +
  ggtitle(expression(bold("Response to N")), subtitle = "CN-model vs.\n MESI + NutNet data") +
  theme_classic() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text( hjust = 0.98, 
                               vjust = 0, 
                               face = 'bold', 
                               margin = margin(t=20, b=-30),
                               size = 17
                               ),
    plot.subtitle = element_text( hjust = 0.98, 
                                  vjust = -3, 
                                  margin = margin(t=20, b=-30),
                                  size = 15
                                  )
  )

gg_mesi_cnmodel_nfert_mesi
ggsave("../fig/gg_mesi_cnmodel_nfert_mesi.pdf")
```

Visualisation of variables from Liang et al.
```{r}
var_order <- c("asat", 
               "vcmax",
               "narea",
               "nmass",
               "lai",
               "leaf_biomass"
               )

# mylabl = rev(c(
#     expression(paste(italic(V)[cmax])),
#     expression(paste(italic(N)[area])),
#     expression(paste(italic(N)[mass])),
#     "LAI",
#     "Leaf biomass"
#   ))

mylabs <- c(
  expression(paste(italic("A")[sat])),
  expression(paste(italic("V")[cmax])),
  expression(paste(italic("N")[area])),
  expression(paste(italic("N")[mass])),
  "LAI",
  # "Leaf area",
  "Leaf biomass"
)

gg_mesi_cnmodel_nfert_liang <- df_box_nfert_liang |> 
  left_join(df_cnmodel_nfert_logresponse, 
            by = "var") |> 
  
  ## give it a nice order (for plotting)
  filter(var %in% var_order) |> 
  mutate(var = factor(var, levels = rev(var_order))) |>  
  
  ggplot() +
  geom_crossbar(aes(x = var, y = middle, ymin = ymin, ymax = ymax), 
                 fill = "#777055ff", 
                color = "black", 
                alpha = 0.6, 
                width = 0.5 
                 ) +
  geom_segment(aes(x = var, xend = var, y = ymin, yend = logr_min)) +
  geom_segment(aes(x = var, xend = var, y = ymax, yend = logr_max)) +
  geom_point(aes(x = var, y = logrr_cnmodel), 
             size = 5, 
             shape = 21, 
             color = "black", 
             fill = "#29a274ff" ) +
  geom_hline( yintercept = 0.0, 
              size = 0.5, 
              linetype = "dotted") +
  ylim(-1, 3.5) +
  scale_x_discrete("", labels = rev(mylabs)) +
  labs(x = "", 
       y = "Log response ratio") +
  ggtitle(expression(bold("Response to N")), 
          subtitle = "CN-model vs.\n Liang et al. data") +
  theme_classic() +
  coord_flip() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text( hjust = 0.98, 
                               vjust = 0, 
                               face = 'bold', 
                               margin = margin(t=20, b=-30),
                               size = 17
                               ),
    plot.subtitle = element_text( hjust = 0.98, 
                                  vjust = -3, 
                                  margin = margin(t=20, b=-30),
                                  size = 15
                                  )
  )

gg_mesi_cnmodel_nfert_liang
ggsave("../fig/gg_mesi_cnmodel_nfert_liang.pdf")
```

## Publication figure

```{r}
# combined meta-analysis plot
cowplot::plot_grid(gg_mesi_cnmodel_nfert_liang,
                   gg_mesi_cnmodel_nfert_mesi,
                   gg_mesi_cnmodel_eco2, 
                   rel_heights = c(0.40, 0.50, 0.77), 
                   labels = c("a", "b", "c"),
                   ncol = 1)

ggsave(paste0(here::here(), "/fig/combined_metaanalysis_cnmodel.pdf"), width = 8, height = 11)
ggsave(paste0(here::here(), "/fig/combined_metaanalysis_cnmodel.png"), width = 8, height = 11)

# cowplot::plot_grid(gg_mesi_cnmodel_eco2, 
#                    gg_mesi_cnmodel_nfert_mesi, 
#                    ncol = 1, 
#                    rel_heights = c(1, 0.43), 
#                    labels = c("a", "b"))
# ggsave(paste0(here::here(), "/fig/mesi_cnmodel.pdf"), width = 6, height = 8)
# ggsave(paste0(here::here(), "/fig/mesi_cnmodel.png"), width = 6, height = 8)

df_correct |> 
  summarise(frac_all = sum(correct_all),
            frac_ci = sum(correct_ci))
```