---
title: "Simulations with cnmodel"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    #theme: paper
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

```{r, warning=FALSE}
library(rsofun)      # make sure to build the cnmodel branch!
library(dplyr)   # for working with ease
library(tidyr)   # for working with ease
library(metafor)     # for meta analysis
library(MAd)         # for meta analysis
library(patchwork)
library(lubridate)
```


# CO2 manipulation

## Model parameters

```{r message=FALSE}
pars <- list(

  # P-model
  kphio                 = 0.04607080,
  soilm_par_a           = 2.75687824,
  soilm_par_b           = 1.68140444,
  tau_acclim_tempstress = 7.35259044,
  par_shape_tempstress  = 0.09863961,

  # Plant
  f_nretain             = 0.500000,
  fpc_tree_max          = 0.950000,
  growtheff             = 0.600000,
  r_root                = 2*0.913000,
  r_sapw                = 2*0.044000,
  exurate               = 0.003000,
  
  k_decay_leaf_base     = 1.00000,
  k_decay_leaf_width    = 2.00000,
  k_decay_root          = 1.00000,
  k_decay_labl          = 0.00000,
  k_decay_sapw          = 1.00000,
  
  r_cton_root           = 37.0000,
  r_cton_wood           = 100.000,
  ncw_min               = 0.056,
  r_n_cw_v              = 0.4,
  r_ctostructn_leaf     = 35.0000,
  kbeer                 = 0.500000,

  # Phenology (should be PFT-specific)
  gddbase               = 5.0,
  ramp                  = 0.0,
  phentype              = 2.0,

  # Soil physics (should be derived from params_soil, fsand, fclay, forg, fgravel)
  perc_k1               = 5.0,        
  thdiff_wp             = 0.2,          
  thdiff_whc15          = 0.8,
  thdiff_fc             = 0.4,          
  forg                  = 0.01,
  wbwp                  = 0.029,  
  por                   = 0.421,    
  fsand                 = 0.82,      
  fclay                 = 0.06,      
  fsilt                 = 0.12,  

  # Water and energy balance
  kA                    = 107,     
  kalb_sw               = 0.17,    
  kalb_vis              = 0.03,    
  kb                    = 0.20,    
  kc                    = 0.25,    
  kCw                   = 1.05,    
  kd                    = 0.50,    
  ke                    = 0.0167,  
  keps                  = 23.44,   
  kWm                   = 220.0,   
  kw                    = 0.26,    
  komega                = 283.0,
  maxmeltrate           = 3.0,

  # Soil BGC
  klitt_af10            = 1.2,
  klitt_as10            = 0.35,
  klitt_bg10            = 0.35,
  kexu10                = 50.0,
  ksoil_fs10            = 0.021,
  ksoil_sl10            = 7.0e-04,
  ntoc_crit1            = 0.45,
  ntoc_crit2            = 0.76,
  cton_microb           = 10.0,
  cton_soil             = 9.77,
  fastfrac              = 0.985,

  # N uptake
  # eff_nup               = 0.600000,  # original value
  eff_nup               = 0.005000,
  minimumcostfix        = 1.000000,
  fixoptimum            = 25.15000,
  a_param_fix           = -3.62000,
  b_param_fix           = 0.270000,

  # Inorganic N transformations
  maxnitr               = 0.1,
  non                   = 0.01,
  n2on                  = 0.0005,
  kn                    = 83.0,
  kdoc                  = 17.0,
  docmax                = 1.0,
  dnitr2n2o             = 0.01,
  
  # Additional parameters - previously forgotten
  beta                  = 146.000000,
  rd_to_vcmax           = 0.01400000,
  tau_acclim            = 10,

  # for development
  tmppar                = 9999

  )
```

## Simulation parameters

```{r}
tmp$params_siml[[1]]$spinupyears <- 2000
tmp$params_siml[[1]]$recycle <- 5
```

## Forcing

Some atmospheric N deposition.
```{r}
tmp <- rsofun::p_model_drivers |> 
  mutate(forcing = purrr::map(forcing, ~mutate(., 
                                               fharv = 0.0,
                                               dno3 = 0.1,
                                               dnh4 = 0.1
                                               )))
```

Harvesting and seed input to zero.
```{r}
use_cseed <- 0 # 100
cn_seed <- 20
use_nseed <- use_cseed / cn_seed

tmp$forcing[[1]] <- tmp$forcing[[1]] |>
  mutate(fharv = ifelse(month(date) == 7 & mday(date) == 15, 0.0, 0.0),
         cseed = ifelse(month(date) == 3 & mday(date) == 15, use_cseed, 0.0),
         nseed = ifelse(month(date) == 3 & mday(date) == 15, use_nseed, 0.0))
```

Synthetic forcing: Constant climate in all days.
```{r}
df_growingseason_mean <- tmp$forcing[[1]] |>
  filter(temp > 5) |>
  summarise(across(where(is.double), .fns = mean))
df_mean <- tmp$forcing[[1]] |>
  summarise(across(where(is.double), .fns = mean))

tmp$forcing[[1]] <- tmp$forcing[[1]] |>
  mutate(temp = df_growingseason_mean$temp,
         prec = df_mean$prec,
         vpd = df_growingseason_mean$vpd,
         ppfd = df_mean$ppfd,
         patm = df_growingseason_mean$patm,
         ccov_int = df_growingseason_mean$ccov_int,
         ccov = df_growingseason_mean$ccov,
         snow = df_mean$snow,
         rain = df_mean$rain,
         fapar = df_mean$fapar,
         co2 = df_growingseason_mean$co2,
         tmin = df_growingseason_mean$tmin,
         tmax = df_growingseason_mean$tmax,
  )
```

Double CO2 from 2010 onwards.
```{r}
tmp$forcing[[1]] <- tmp$forcing[[1]] |>
  mutate(co2 = ifelse(year(date) >= 2010, co2 * 2, co2))
```

Repeat last year's forcing N times to investigate multi-decadal dynamics.
```{r}
n_ext <- 100
for (idx in seq(n_ext)){
  tmp$forcing[[1]] <- bind_rows(
    tmp$forcing[[1]],
    tmp$forcing[[1]] |> 
      tail(365) |> 
      mutate(date = date + years(idx))
  )
}
tmp$params_siml[[1]]$nyeartrend <- tmp$params_siml[[1]]$nyeartrend + n_ext
```


# Model run

```{r}
output <- runread_pmodel_f(
  tmp,
  par = pars
  )

output <- output$data[[1]]
```

# Visualisations

## Response ratios

Rearrange outputs to get response ratios.
```{r}
df_out <- output |> 
  mutate(leaf_cn = cleaf/nleaf, root_shoot_ratio = croot/cleaf, n_inorg = pno3 + pnh4) |> 
  select(date, gpp, vcmax, jmax, gs = gs_accl, leaf_cn, lai, agb = cleaf, 
         bgb = croot, root_shoot_ratio, nup, n_inorg, n_uptake = nup)
  
df_amb <- df_out |> 
  filter(year(date) < 2010) |> 
  summarise(across(where(is.numeric), mean))

df_ele <- df_out |> 
  filter(year(date) %in% 2010:2012) |> 
  summarise(across(where(is.numeric), mean))

df_ele2 <- df_out |> 
  filter(year(date) %in% 2100:2107) |> 
  summarise(across(where(is.numeric), mean))

df_exp <- bind_rows(df_amb, df_ele)
df_rr  <- log(df_exp[2,]/df_exp[1,]) |> 
  pivot_longer(cols = everything(), names_to = "variable", values_to = "response") |> 
  mutate(variable = factor(variable, 
                          levels = rev(c("gpp", "vcmax", "jmax", "gs", "leaf_cn", "lai", "agb", 
                                       "bgb", "root_shoot_ratio", "n_uptake", "n_inorg"))))

df_exp2 <- bind_rows(df_amb, df_ele2)
df_rr2  <- log(df_exp2[2,]/df_exp2[1,]) |> 
  pivot_longer(cols = everything(), names_to = "variable", values_to = "response") |> 
  mutate(variable = factor(variable, 
                           levels = rev(c("gpp", "vcmax", "jmax", "gs", "leaf_cn", "lai", "agb", 
                                       "bgb", "root_shoot_ratio", "n_uptake", "n_inorg"))))
```

Plot model results on top of meta-analysis results from MESI.
```{r}
df_box <- readRDS("../data/df_box_mesi_co2.rds")
ggplot() +
  geom_crossbar( aes(x = var, y = middle, ymin = ymin, ymax = ymax), 
                 data = df_box |> 
                   mutate(var = factor(var, levels = rev(c("asat", "gpp", "vcmax", "jmax", "gs", "leaf_n", 
                                                           "leaf_cn", "lai", "leaf_area", "leaf_biomass", "anpp", "agb", 
                                                           "root_production", "bgb", "root_shoot_ratio", "n_uptake", "n_inorg")))),
                 fill = "tomato", color = "black", alpha = 0.6, width = 0.5 
                 ) +
  geom_point(aes(var, response), 
             data = df_rr2 |> 
               rename(var = variable) |> 
               mutate(var = factor(var, levels = rev(c("asat", "gpp", "vcmax", "jmax", "gs", "leaf_n", 
                                                       "leaf_cn", "lai", "leaf_area", "leaf_biomass", "anpp", "agb", 
                                                       "root_production", "bgb", "root_shoot_ratio", "n_uptake", "n_inorg")))), 
             size = 3, color = "royalblue") +
  geom_hline( yintercept = 0.0, size = 0.5 ) +
  labs(x = "Variable", y = "Log Response Ratio", size = expression(paste("Error"^{-1}))) +
  coord_flip() +
  ylim(-1, 1) +
  labs(title = "cnmodel vs. MESI", subtitle = "Response to eCO2")
```
