---
title: "Terrestrial sink trends"
output: html_document
date: "2022-09-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(cowplot)
library(RColorBrewer)
```

## GCB 2021 / TRENDY v8

Corresponding published article doi: https://doi.org/10.5194/essd-14-1917-2022

Data obtained from: doi:10.18160/gcp-2021

File `data/Global_Carbon_Budget_2021v1.0_tab_terr_sink.csv` is copied from `Global_Carbon_Budget_2021v1.0.xlsx`, tab "Terrestrial Sink" and exported as CSV. 

File `data/Global_Carbon_Budget_2021v1.0_tab_gcb.csv` is copied from `Global_Carbon_Budget_2021v1.0.xlsx`, tab "Global Carbon Budget" and exported as CSV. 

(Internal: see also `~/data/globalcarbonproject/gcp-2021/`.)

```{r cars}
df_models <- read_csv(here::here("data/Global_Carbon_Budget_2021v1.0_tab_terr_sink.csv"))
df_gcb <- read_csv(here::here("data/Global_Carbon_Budget_2021v1.0_tab_gcb.csv"))
```

Get residual sink from the budget, interpreted as the implied land sink. Note that the GCB presents the land sink as the multi-model mean (after accounting for LUC). 
```{r}
df_gcb <- df_gcb |> 
  mutate(s_land_impl = e_ff + e_luc - s_cement - s_ocean - g_atm)

df_gcb |> 
  pivot_longer(cols = c(s_land, s_land_impl), values_to = "s_land", names_to = "source") |> 
  ggplot(aes(x = year, y = s_land, color = source)) +
  geom_line()

# combine
df <- df_models |> 
  left_join(
    df_gcb |> 
      rename(Year = year),
    by = "Year"
  )
```

Define decades
```{r pressure, echo=FALSE}
df <- df |> 
  mutate(decade = ifelse(Year > 1979 & Year < 1990, 
                         1980,
                         ifelse(Year > 1989 & Year < 2000, 1990,
                                ifelse(Year > 1999 & Year < 2010, 2000,
                                       ifelse(Year > 2009, 2010, NA)))
                         ))
```

Identify models as C-only and C-N models. The table loaded here (`data/model_types_trendy_gcp2021.csv`) is created manually based on information of the corresponding Global Carbon Budget paper (in this case GCB 2021, Friedlingstein et al., 2022 https://essd.copernicus.org/articles/14/1917/2022/) using their Table A.1.

```{r}
# data frame in long format
df_models <- df |> 
  select(-GCB, -mean_models, -sd_models) |> 
  pivot_longer(cols = c("CABLE-POP",  
                        "CLASSIC",    
                        "CLM5.0",     
                        "DLEM",       
                        "IBIS",       
                        "ISAM",       
                        "ISBA-CTRIP", 
                        "JSBACH",     
                        "JULES-ES", 
                        "LPJ-GUESS",  
                        "LPJ",        
                        "LPX-Bern",   
                        "OCNv2",      
                        "ORCHIDEE-v3",
                        "SDGVM",      
                        "VISIT",      
                        "YIBs"),
               names_to = "model",
               values_to = "sland"
              )

# get information on which model is C-only and which model is C-N from manually 
# created table, based on Table A1 in Friedlingstein et al. 2021
df_modeltypes <- read_csv(here::here("data/model_types_trendy_gcp2021.csv"))

# aggregate (mean) by decade and model
df_decades <- df_models |> 
  drop_na() |> 
  group_by(decade, model) |> 
  summarise(sland = mean(sland)) |> 
  left_join(
    df_modeltypes,
    by = "model"
  )

# add info on model type to un-aggregated (containing all years) data frame
df_models <- df_models |> 
  left_join(
    df_modeltypes,
    by = "model"
  )

# aggregate observations (implied land sink) by decade
df_decades_obs <- df |> 
  drop_na() |> 
  group_by(decade) |> 
  summarise(sland = mean(s_land_impl))
```
## Exploratory plot

```{r}
ggplot() +
  geom_boxplot(data = df_decades, aes(as.factor(decade), sland, fill = type)) +
  geom_jitter(data = df_decades, aes(as.factor(decade), sland, color = type), width = 0.1) +
  geom_point(data = df_decades_obs, aes(as.factor(decade), sland), color = "black", size = 5)
```

## Publication figure
```{r}
# re-arrange data for plotting
tmp <- df_models |> 
  
  # aggregate by model type (mean)
  group_by(Year, type) |> 
  summarise(sland = mean(sland)) |> 

  # add aggregated by model type (min)
  left_join(
    df_models |> 
      group_by(Year, type) |> 
      summarise(sland_min = min(sland)) 
  ) |> 

  # add aggregated by model type (max)
  left_join(
    df_models |> 
      group_by(Year, type) |> 
      summarise(sland_max = max(sland)) 
  ) |> 

  # add implied residual sink as "Obs."
  bind_rows(
    df |> 
      select(Year, sland = s_land_impl) |> 
      mutate(type = "Obs.")
  )

# make model type a factor
vec_types <- unique(tmp$type)
tmp <- tmp |> 
  mutate(type = factor(type, 
                        levels = (c("Obs.", 
                                    vec_types[-which(vec_types == "Obs.")]))))

# create colour palette distinguishing models
colourCount = length(unique(tmp$type))
getPalette = colorRampPalette(brewer.pal(12, "Paired"))
set.seed(1982)

# time series
gg1 <- tmp |> 
  ggplot(aes(Year, sland, color = type)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_ribbon(aes(x = Year, ymin = sland_min, ymax = sland_max, fill = type), color = NA) +
  geom_line(linewidth = 0.8) +
  theme_classic() +
  scale_color_manual(values = c("black", "tomato", "royalblue"), 
                     name = "") +
  scale_fill_manual(values = c(rbeni::add_alpha("black", 0.3), 
                               rbeni::add_alpha("tomato", 0.3), 
                               rbeni::add_alpha("royalblue", 0.3)),
                    name = "") +
  labs(x = "Year", y = expression(paste("Sink (Pg C yr"^-1, ")")))

get_trend <- function(adf, year_start){
  adf <- adf |> 
    filter(Year >= year_start)
  linmod <- lm(sland ~ Year, data = adf)
  sinktrend <- coef(linmod)["Year"]
  return(sinktrend)
}

get_trend_unc <- function(adf, year_start){
  adf <- adf |> 
    filter(Year >= year_start)
  linmod <- lm(sland ~ Year, data = adf)
  unc <- summary(linmod)$coefficients["Year", "Std. Error"]
  return(unc)
}

get_iav <- function(adf, year_start){
  adf <- adf |> 
    filter(Year >= year_start)
  linmod <- lm(sland ~ Year, data = adf)
  iav <- sd(linmod$residuals)
  return(iav)
}

get_sland_mean <- function(adf, use_years){
  adf |> 
    filter(Year %in% use_years) |> 
    summarise(sland = mean(sland)) |> 
    pull(sland)
}

get_sland_sd <- function(adf, use_years){
  adf |> 
    filter(Year %in% use_years) |> 
    summarise(sland = sd(sland)) |> 
    pull(sland)
}

df_models_trend <- df_models |> 
  bind_rows(
    df |> 
      select(Year, sland = s_land_impl) |> 
      mutate(model = "obs")
  ) |> 
  group_by(model) |> 
  nest() |> 
  mutate(
    trend = purrr::map_dbl(data, ~get_trend(., year_start = 1959)),
    trend_unc = purrr::map_dbl(data, ~get_trend_unc(., year_start = 1959)),
    iav = purrr::map_dbl(data, ~get_iav(., year_start = 1959)),
    sland_mean = purrr::map_dbl(data, ~get_sland_mean(., use_years = 2011:2020)),
    sland_sd = purrr::map_dbl(data, ~get_sland_sd(., use_years = 2011:2020))
    ) |> 
  left_join(
    df_modeltypes,
    by = "model"
  ) |> 
  mutate(type = ifelse(is.na(type), "obs", type)) |> 
  mutate(model = ifelse(model == "obs", "Observations", model)) |> 
  mutate(type = ifelse(type == "obs", "Obs.", type))

colourCount = length(unique(df_models_trend$model))
getPalette = colorRampPalette(brewer.pal(12, "Paired"))
set.seed(1986)

df_plot2 <- df_models_trend |> 
  mutate(type = factor(type, levels = c("CN", "C", "Obs.")),
         model = factor(model, 
                        levels = (c("Observations", 
                                    df_models_trend$model[-which(df_models_trend$model == "Observations")]))))

gg2 <- ggplot() +
  geom_boxplot(
    aes(
      x = type, 
      y = sland_mean
      ),
    data = df_plot2 |> 
      mutate(sland_mean = ifelse(type == "Obs.", NA, sland_mean)),
    fill = "azure3", 
    outlier.color = "white"
    ) +
  geom_jitter(
    aes(x = type, y = sland_mean, color = model),
    data = df_plot2 |> 
      mutate(sland_mean = ifelse(type == "Obs.", NA, sland_mean)),
    width = 0.2, 
    size = 3, 
    show.legend = FALSE
    ) + 
  geom_pointrange(
    aes(
      x = type, 
      y = sland_mean, 
      color = model, 
      ymin = sland_mean - 1.96 * sland_sd, 
      ymax = sland_mean + 1.96 * sland_sd
      ),
    data = df_plot2 |> 
      mutate(sland_mean = ifelse(type != "Obs.", NA, sland_mean)),
    width = 0.2, 
    size = 1, 
    show.legend = FALSE
    ) + 
  theme_classic() + 
  scale_color_manual(
    values = c(
      "CABLE-POP" = getPalette(colourCount)[1],
      "CLASSIC" = getPalette(colourCount)[2],
      "CLM5.0" = getPalette(colourCount)[3],
      "DLEM" = getPalette(colourCount)[4],
      "IBIS" = getPalette(colourCount)[5],
      "ISAM" = getPalette(colourCount)[6],
      "ISBA-CTRIP" = getPalette(colourCount)[7],
      "JSBACH" = getPalette(colourCount)[8],
      "JULES-ES" = getPalette(colourCount)[9],
      "LPJ-GUESS" = getPalette(colourCount)[10],
      "LPJ" = getPalette(colourCount)[11],
      "LPX-Bern" = getPalette(colourCount)[12],
      "OCNv2" = getPalette(colourCount)[13],
      "ORCHIDEE-v3" = getPalette(colourCount)[14],
      "SDGVM" = getPalette(colourCount)[15],
      "VISIT" = getPalette(colourCount)[16],
      "YIBs" = getPalette(colourCount)[17],
      "Observations" = "black"
    )
    ) +
  labs(x = "Model type and observations", 
       y = expression(paste("Sink (Pg C yr"^-1, ")")),
       title = "Mean 2011-2020") +
  theme(
    legend.title = element_blank()
  )

df_plot3 <- df_models_trend |> 
  mutate(type = factor(type, levels = c("CN", "C", "Obs.")),
         model = factor(model, 
                        levels = (c("Observations", 
                                    df_models_trend$model[-which(df_models_trend$model == "Observations")]))))

gg3 <- ggplot() +
  geom_boxplot(
    aes(
      x = type, 
      y = trend
      ),
    data = df_plot3 |> 
      mutate(trend = ifelse(type == "Obs.", NA, trend)),
    fill = "azure3", 
    outlier.color = "white"
    ) +
  geom_jitter(
    aes(
      x = type, 
      y = trend, 
      color = model
      ),
    data = df_plot3 |> 
      mutate(trend = ifelse(type == "Obs.", NA, trend)),
    width = 0.2, 
    size = 3
    ) + 
  geom_pointrange(
    aes(
      x = type, 
      y = trend, 
      color = model, 
      ymin = trend - 1.96 * trend_unc, 
      ymax = trend + 1.96 * trend_unc
      ),
    data = df_plot2 |> 
      mutate(trend = ifelse(type != "Obs.", NA, trend)),
    width = 0.2, 
    size = 1, 
    show.legend = FALSE
    ) + 
  theme_classic() + 
  scale_color_manual(
    values = c(
      "CABLE-POP" = getPalette(colourCount)[1],
      "CLASSIC" = getPalette(colourCount)[2],
      "CLM5.0" = getPalette(colourCount)[3],
      "DLEM" = getPalette(colourCount)[4],
      "IBIS" = getPalette(colourCount)[5],
      "ISAM" = getPalette(colourCount)[6],
      "ISBA-CTRIP" = getPalette(colourCount)[7],
      "JSBACH" = getPalette(colourCount)[8],
      "JULES-ES" = getPalette(colourCount)[9],
      "LPJ-GUESS" = getPalette(colourCount)[10],
      "LPJ" = getPalette(colourCount)[11],
      "LPX-Bern" = getPalette(colourCount)[12],
      "OCNv2" = getPalette(colourCount)[13],
      "ORCHIDEE-v3" = getPalette(colourCount)[14],
      "SDGVM" = getPalette(colourCount)[15],
      "VISIT" = getPalette(colourCount)[16],
      "YIBs" = getPalette(colourCount)[17],
      "Observations" = "black"
    )
    ) +
  labs(x = "Model type and observations", 
       y = expression(paste("Sink trend (Pg C yr"^-2, ")")),
       title = "Trend 1959-2020") +
  theme(
    legend.title = element_blank()
  )


bottomrow <- cowplot::plot_grid(gg2, gg3, nrow = 1, rel_widths = c(0.8, 1), labels = c("b", "c"))
cowplot::plot_grid(gg1, bottomrow, nrow = 2, rel_heights = c(0.8, 1), labels = c("a", "", ""))
ggsave(here::here("fig_pub/sink_trend_trendy.pdf"), width = 1.7*6, height = 1.7*5)
ggsave(here::here("fig_pub/sink_trend_trendy.png"), width = 1.7*6, height = 1.7*5)

gg1
ggsave(here::here("fig/sink_trend_trendy_tseries.pdf"), width = 1.7*6, height = 4)

gg3
ggsave(here::here("fig/sink_trend_trendy_boxplot_trend.pdf"), width = 6, height = 5)
```
